### 2024-11-29 paper/paper.qmd
more absolutely sicko cleaning

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=1810000402

https://data.ontario.ca/dataset/school-board-achievements-and-progress/resource/9d81dc27-32ef-4864-94b5-f09950d00c72

https://data.ontario.ca/dataset/school-board-financial-reports-estimates-revised-estimates-and-financial-statements/resource/56598892-c180-4f0d-bde5-63f971d6ef9a

https://www150.statcan.gc.ca/n1/en/catalogue/81-582-X

### 2024-11-29 scripts/02-data_cleaning.R

@@ -38,38 +38,58 @@ colnames(data_values) <- c("Board Number", "Total Expenses", "Expenditure on Fac
 merged_data <- cleaned_data %>%
   left_join(data_values, by = "Board Number")
 
-# View the final merged dataset
-head(merged_data)
 
 
+table3 <- read_excel(here::here("./data/raw_data/education_indicators.xlsx"))
 
+relevant_data <- table3 %>%
+  select(board_number = 1, 
+         enrolment = 22, 
+         percent_low_income = 47, 
+         percent_no_degree = 48)
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+relevant_data <- relevant_data %>%
   mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+    enrolment = as.numeric(enrolment),
+    percent_low_income = as.numeric(percent_low_income),
+    percent_no_degree = as.numeric(percent_no_degree)
+  )
 
-#### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+# Filter out rows where numeric columns have NA (invalid conversion)
+relevant_data <- relevant_data %>%
+  filter(!is.na(enrolment), !is.na(percent_low_income), !is.na(percent_no_degree))
+
+
+# Group by board_number and calculate weighted averages for percentages
+aggregated_data <- relevant_data %>%
+  group_by(board_number) %>%
[…]

### 2024-11-30 scripts/01-download_data.R
updated download_data script to download data and updated cleaning to work with the new data + bugfixes

@@ -7,20 +7,69 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
+library(httr)
+library(here)
 
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
+# Define the URL and local file path
+url <- "https://data.ontario.ca/dataset/71244894-f3e1-4be0-a69a-9bdc4eb45ac1/resource/9d81dc27-32ef-4864-94b5-f09950d00c72/download/sbpr_data_table_en_sept2024.xlsx"
+referrer <- "https://data.ontario.ca/dataset/71244894-f3e1-4be0-a69a-9bdc4eb45ac1"
+local_file <- here::here("data", "raw_data", "academic_performance.xlsx")
+# Use httr with Referer and User-Agent headers
+response <- GET(
+  url,
+  add_headers(
+    `Referer` = referrer,
+    `User-Agent` = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/131.0.0.0 Safari/537.36"
+  )
+)
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# Check if the request was successful
+if (status_code(response) == 200) {
+  writeBin(content(response, "raw"), local_file)
+  message("File successfully downloaded to: ", local_file)
+} else {
+  message("Failed to download the file. HTTP status: ", status_code(response))
+}
 
 
+url <- "https://data.ontario.ca/dataset/f9393864-17ae-43f4-a0e1-062fddc6c99c/resource/56598892-c180-4f0d-bde5-63f971d6ef9a/download/efis_2024-25_estimates_-_board_submitted.csv"
+referrer <- "https://data.ontario.ca/dataset/f9393864-17ae-43f4-a0e1-062fddc6c99c"
+local_file <- here::here("data", "raw_data", "school_expenditures.csv")
 
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+# Use httr with Referer and User-Agent headers
+response <- GET(
+  url,
+  add_headers(
+    `Referer` = referrer,
[…]

### 2024-11-30 scripts/02-data_cleaning.R

@@ -24,8 +24,8 @@ cleaned_data <- raw_data %>%
 raw_data_2 <- read_csv(here::here("./data/raw_data/school_expenditures.csv"))
 
 cleaned_data_2 <- raw_data_2 %>%
-  slice(c(1, 16, 47)) %>%  # Select specific rows
-  select(10:ncol(raw_data_2))       # Select columns from the 10th onward
+  slice(c(1, 16, 46)) %>%  # Select specific rows
+  select(9:ncol(raw_data_2))       # Select columns from the 10th onward
 
 data_values <- cleaned_data_2[1:3, ]
 # Transpose data_values to align with board numbers

### 2024-12-02 paper/paper.qmd
updated download_data script to download data and updated cleaning to work with the new data + bugfixes
everything

@@
 
-https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=1810000402
-
 https://data.ontario.ca/dataset/school-board-achievements-and-progress/resource/9d81dc27-32ef-4864-94b5-f09950d00c72
@@
 
-https://www150.statcan.gc.ca/n1/en/catalogue/81-582-X
+https://data.ontario.ca/dataset/school-information-and-student-demographics/resource/e0e90bd5-d662-401a-a6d2-60d69ac89d14
+
+https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1810000501
+
+https://stats.stackexchange.com/questions/11676/pseudo-r-squared-formula-for-glms

### 2024-12-02 scripts/00-simulate_data.R

@@ -1,19 +1,46 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates data from the final analysis dataset
+# Author: Andrew Yu Xuan Goh
+# Date: 1 December 2024
+# Contact: andrew.goh@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
+# Pre-requisites: None
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(arrow)
+
+#### Define simulation parameters ####
+# Set random seed for reproducibility
+set.seed(42)
+
+# Define the number of rows to simulate
+n_rows <- 100
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# Generate simulated data for each column
+simulated_data <- tibble(
+  `Board Number` = sprintf("B%03d", 1:n_rows),  # Simulated board numbers (e.g., B001, B002)
+  `Board Name` = paste("Board", 1:n_rows),     # Simulated board names (e.g., Board 1, Board 2)
+  Region = sample(c("Region A", "Region B", "Region C"), n_rows, replace = TRUE),  # Random regions
+  City = sample(c("City X", "City Y", "City Z"), n_rows, replace = TRUE),          # Random cities
+  `Four Year Graduation Rate` = round(runif(n_rows, 0.7, 1), 3),                   # Graduation rates between 0.7 and 1
+  `Total Expenses` = round(runif(n_rows, 1e6, 1e8), 3),                            # Total expenses between $1M and $100M
+  `Expenditure on Facilities` = round(runif(n_rows, 1e5, 1e7), 3),                 # Facility expenses between $100K and $10M
+  `Total Enrolment` = round(runif(n_rows, 100, 5000)),                             # Total enrolment between 100 and 5000 students
+  percent_low_income = round(runif(n_rows, 10, 50), 3),                            # Low-income percentage between 10% and 50%
+  percent_no_degree = round(runif(n_rows, 5, 30), 3)                               # No-degree percentage between 5% and 30%
+) %>%
[…]

### 2024-12-02 scripts/01-test_simulated_data.R

@@ -1,0 +1,121 @@
+#### Preamble ####
+# Purpose: Tests the simulated dataset.
+# Author: Andrew Yu Xuan Goh
+# Date: 1 December 2024
+# Contact: andrew.goh@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Has the simulated data in the form of a parquet.
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+library(testthat)
+
+#### Load data ####
+# Load simulated data from the parquet file
+simulated_data <- read_parquet(here::here("./data/simulated_data/simulated_data.parquet"))
+
+#### Define sanity check function ####
+# Function to perform tests on the dataset
+run_sanity_checks <- function(data, dataset_name) {
+  test_that(paste("Sanity Checks for", dataset_name), {
+    
+    #### Data structure tests ####
+    # Check if the dataset is a dataframe or tibble
+    expect_true(is.data.frame(data), info = paste(dataset_name, "is not a dataframe or tibble"))
+    
+    # Verify that the dataset has an appropriate number of rows
+    expect_true(nrow(data) >= 5 && nrow(data) <= 1000,
+                info = paste(dataset_name, "does not have between 5 and 1000 rows")
+    )
+    
+    # Ensure the dataset has exactly 11 columns
+    expect_equal(ncol(data), 11,
+                 info = paste(dataset_name, "does not have exactly 11 columns")
+    )
+    
+    # Confirm all required columns are present
+    required_cols <- c(
+      "Board Number", "Board Name", "Region", "City",
+      "Four Year Graduation Rate", "Total Expenses",
+      "Expenditure on Facilities", "Total Enrolment",
+      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities"
+    )
+    expect_true(all(required_cols %in% colnames(data)),
+                info = paste(dataset_name, "is missing required columns")
+    )
+    
+    #### Data type tests ####
+    # Check the types of key variables
+    expect_type(data$`Board Number`, "character")
+    expect_type(data$`Board Name`, "character")
+    expect_type(data$Region, "character")
+    expect_type(data$City, "character")
+    expect_type(data$`Four Year Graduation Rate`, "double")
+    expect_type(data$`Total Expenses`, "double")
+    expect_type(data$`Expenditure on Facilities`, "double")
+    expect_type(data$`Total Enrolment`, "double")
+    expect_type(data$percent_low_income, "double")
+    expect_type(data$percent_no_degree, "double")
[…]

### 2024-12-02 scripts/02-download_data.R

@@ -1,15 +1,15 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from data.ontario.ca
+# Author: Andrew Yu Xuan Goh
+# Date: 30 November 2024
+# Contact: andrew.goh@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: None
+#### Workspace setup ####
 library(httr)
 library(here)
 
+#### Download data ####
 # Define the URL and local file path
 url <- "https://data.ontario.ca/dataset/71244894-f3e1-4be0-a69a-9bdc4eb45ac1/resource/9d81dc27-32ef-4864-94b5-f09950d00c72/download/sbpr_data_table_en_sept2024.xlsx"
 referrer <- "https://data.ontario.ca/dataset/71244894-f3e1-4be0-a69a-9bdc4eb45ac1"

### 2024-12-02 scripts/03-clean_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
-# Purpose: Cleans the raw data recorded from dataontario and statcan
-# Author: Andrew Goh Yu Xuan
+# Purpose: Cleans the raw data recorded from data.ontario.ca and saves it as a parquet
+# Author: Andrew Yu Xuan Goh
 # Date: 27 November 2024
 # Contact: andrew.goh@mail.utoronto.ca
 # License: MIT
@@ -9,9 +9,13 @@
 library(dplyr)
 library(tidyverse)
 library(readxl)
+library(arrow)
+
 #### Clean data ####
+# Read academic performance data
 raw_data <- read_excel(here::here("./data/raw_data/academic_performance.xlsx"))
 
+# Remove unused variables in the data
 cleaned_data <- raw_data %>%
   select(-contains("Progress"), 
          -contains("Credit Accumulation"), 
@@ -19,35 +23,36 @@ cleaned_data <- raw_data %>%
          -contains("Language"),
          -contains("Type"),
          -contains("Five")) %>%
-  drop_na()  
+  drop_na()  # Cleans out N/A values
 
+# Read school expenditures data
 raw_data_2 <- read_csv(here::here("./data/raw_data/school_expenditures.csv"))
 
+# Remove unused variables in the data (as well as filler rows/columns in this case)
 cleaned_data_2 <- raw_data_2 %>%
   slice(c(1, 16, 46)) %>%  # Select specific rows
   select(9:ncol(raw_data_2))       # Select columns from the 10th onward
 
-data_values <- cleaned_data_2[1:3, ]
-# Transpose data_values to align with board numbers
-data_values <- t(data_values)
-data_values <- as.data.frame(data_values)
-colnames(data_values) <- c("Board Number", "Total Expenses", "Expenditure on Facilities")  # Assign new column names
+# Transpose data to be in the same format as the first table
+cleaned_data_2 <- t(cleaned_data_2)
+cleaned_data_2 <- as.data.frame(cleaned_data_2)
+colnames(cleaned_data_2) <- c("Board Number", "Total Expenses", "Expenditure on Facilities")  # Assign new column names
 
-
-# Merge the data based on board_number
+# Merge the data from the two tables based on board_number
 merged_data <- cleaned_data %>%
-  left_join(data_values, by = "Board Number")
+  left_join(cleaned_data_2, by = "Board Number")
 
-
-
+# Read education indicators data
[…]

### 2024-12-02 scripts/04-test_analysis_data.R

@@ -1,0 +1,121 @@
+#### Preamble ####
+# Purpose: Tests the analysis dataset.
+# Author: Andrew Yu Xuan Goh
+# Date: 1 December 2024
+# Contact: andrew.goh@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Has the analysis data in the form of a parquet.
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+library(testthat)
+
+#### Load data ####
+# Load analysis data from the parquet file
+analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.parquet"))
+
+#### Define sanity check function ####
+# Function to perform tests on the dataset
+run_sanity_checks <- function(data, dataset_name) {
+  test_that(paste("Sanity Checks for", dataset_name), {
+    
+    #### Data structure tests ####
+    # Check if the dataset is a dataframe or tibble
+    expect_true(is.data.frame(data), info = paste(dataset_name, "is not a dataframe or tibble"))
+    
+    # Verify that the dataset has an appropriate number of rows
+    expect_true(nrow(data) >= 5 && nrow(data) <= 1000,
+                info = paste(dataset_name, "does not have between 5 and 1000 rows")
+    )
+    
+    # Ensure the dataset has exactly 11 columns
+    expect_equal(ncol(data), 11,
+                 info = paste(dataset_name, "does not have exactly 11 columns")
+    )
+    
+    # Confirm all required columns are present
+    required_cols <- c(
+      "Board Number", "Board Name", "Region", "City",
+      "Four Year Graduation Rate", "Total Expenses",
+      "Expenditure on Facilities", "Total Enrolment",
+      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities"
+    )
+    expect_true(all(required_cols %in% colnames(data)),
+                info = paste(dataset_name, "is missing required columns")
+    )
+    
+    #### Data type tests ####
+    # Check the types of key variables
+    expect_type(data$`Board Number`, "character")
+    expect_type(data$`Board Name`, "character")
+    expect_type(data$Region, "character")
+    expect_type(data$City, "character")
+    expect_type(data$`Four Year Graduation Rate`, "double")
+    expect_type(data$`Total Expenses`, "double")
+    expect_type(data$`Expenditure on Facilities`, "double")
+    expect_type(data$`Total Enrolment`, "double")
+    expect_type(data$percent_low_income, "double")
+    expect_type(data$percent_no_degree, "double")
[…]

### 2024-12-02 scripts/05-model.R

@@ -1,0 +1,113 @@
+#### Preamble ####
+# Purpose: Models the relationships between variables in the data.
+# Author: Andrew Yu Xuan Goh
+# Date: 1 November 2024
+# Contact: andrew.goh@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Has analysis data saved as a parquet
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+library(brms)
+library(broom)
+library(caret)
+
+#### Load data ####
+# Read analysis data from the parquet file
+analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.parquet"))
+
+# Rename columns to remove spaces
+colnames(analysis_data) <- gsub(" ", "_", colnames(analysis_data))
+
+# Check the new column names
+colnames(analysis_data)
+
+#### Data Preprocessing ####
+# Select relevant columns for modeling
+data <- analysis_data %>%
+  dplyr::select(Four_Year_Graduation_Rate, 
+         Total_Expenses, 
+         percent_no_degree, 
+         percent_low_income, 
+         percentage_spent_on_facilities)
+
+# Normalize data using caret package for scaling
+data_scaled <- data %>%
+  mutate(across(everything(), ~ scale(.) %>% as.vector()))
+
+#### Linear Regression Model ####
+# Fit the linear regression model
+lm_model <- lm(Four_Year_Graduation_Rate ~ Total_Expenses + percent_no_degree + percent_low_income + percentage_spent_on_facilities, data = data_scaled)
+
+# Print model summary
+lm_summary <- summary(lm_model)
+lm_stats <- broom::tidy(lm_model)
+
+# Print R-squared and p-values
+print("Linear Model Summary:")
+print(lm_summary)
+
+# Calculate R-squared
+rsq <- lm_summary$r.squared
+print(paste("R-squared for Linear Regression: ", round(rsq, 3)))
+
+# Print p-values for each predictor
+print("P-values for each predictor in the linear regression model:")
+print(lm_stats)
+
+#### Bayesian Linear Regression Model ####
+# Fit the Bayesian linear regression model using the brms package
+bayesian_model <- brm(
+  Four_Year_Graduation_Rate ~ Total_Expenses + percent_no_degree + percent_low_income + percentage_spent_on_facilities, 
+  data = data_scaled,
+  family = gaussian(),
+  iter = 2000, 
+  warmup = 1000, 
[…]

### 2024-12-04 paper/paper.qmd
models

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+title: "Facilities Spending and Socioeconomic Factors Show Strong Correlation with Ontario School District Graduation Rates"
+author: Andrew Goh
+thanks: "Code and data are available at: https://github.com/mushroomcarbon/schools_and_money"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Using data from Ontario's school board achievement, financial, and demographic datasets, we analyze the relationship between various district-level factors and four-year graduation rates.
+Through polynomial regression analysis, we find that the percentage of households without post-secondary degrees and facilities spending have significant negative correlations with graduation rates, while per-student expenses show a non-linear relationship.
+These findings suggest that socioeconomic factors and resource allocation decisions significantly impact academic outcomes, with implications for education policy and resource distribution in Ontario school districts."
 format: pdf
+header-includes:
+  - \usepackage{float}
+  - \usepackage{placeins}
 number-sections: true
@@
 library(tidyverse)
-library(palmerpenguins)
+library(arrow)
+library(broom)
+library(gt)
+library(here)
+library(car)
+library(caret)
+
+# Read the analysis data
+analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.parquet"))
+
+# Create summary statistics
+summary_stats <- analysis_data %>%
+  summarise(across(c(`Four Year Graduation Rate`, 
+                     `Total Expenses`, 
+                     `Total Enrolment`,
+                     percentage_spent_on_facilities,
+                     expenses_per_quota,
+                     percent_no_degree,
+                     percent_low_income),
+                   list(
+                     Mean = ~mean(., na.rm = TRUE),
+                     SD = ~sd(., na.rm = TRUE),
+                     Min = ~min(., na.rm = TRUE),
+                     Max = ~max(., na.rm = TRUE),
+                     Median = ~median(., na.rm = TRUE)
+                   )))
+
+# Create scaled version for modeling
+data_scaled <- analysis_data %>%
+  mutate(across(where(is.numeric), ~ scale(.) %>% as.vector()))
+
+# Load the polynomial model from the RDS file
+poly_model <- readRDS(here("models", "poly_model.rds"))
+lm_model <- readRDS(here("models", "lm_model.rds"))
+
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+The relationship between educational resources and academic outcomes has been a central focus of education policy research, particularly in publicly funded school systems.
+In Ontario, where school boards receive varying levels of funding and serve communities with diverse socioeconomic profiles, understanding how different factors influence student success is crucial for effective policy-making [@ontarioeducation].
 
-The remainder of this paper is structured as follows. @sec-data....
+Recent studies have suggested that both financial resources and community characteristics play significant roles in determining student outcomes [@educationresources; @socioeconomic].
+However, the specific mechanisms through which these factors interact, particularly in the Ontario context, remain unclear.
+This gap in understanding is particularly relevant given Ontario's significant investment in education, with annual spending exceeding $31 billion [@ontariobudget].
+
+Our paper examines how various district-level factors—including financial allocations, socioeconomic indicators, and operational characteristics—correlate with four-year graduation rates across Ontario school boards.
+We focus specifically on the relationships between graduation rates and six key variables: total expenses, percentage of households without post-secondary degrees, percentage of low-income households, percentage of budget spent on facilities, per-student expenses, and total enrollment.
+
+Our estimand is the expected change in four-year graduation rates associated with changes in these district-level characteristics, particularly focusing on the non-linear relationships revealed through polynomial regression analysis.
+
+The results suggest that socioeconomic factors, particularly the percentage of households without post-secondary degrees, have a significant negative linear relationship with graduation rates.
+Additionally, we find that the relationship between per-student expenses and graduation rates follows a U-shaped curve, suggesting diminishing returns up to a certain point.
+These findings have important implications for how educational resources might be most effectively allocated to improve student outcomes.
+
+The remainder of this paper is organized as follows: @sec-data describes our dataset and measurement approach, @sec-model outlines our polynomial regression methodology, @sec-results presents our findings, and @sec-discussion explores the implications and limitations of our analysis.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Overview
+We use the statistical programming language R [@citer] for the analysis and presentation of the project.
+The tidyverse [@tidyverse] ecosystem, particularly dplyr [@dplyr], was used for data manipulation.
+For model development and evaluation, we utilized caret [@caret] for model training and glmnet [@glmnet1; @glmnet2; @glmnet3] for regularized regression methods.
+The arrow [@arrow] package was employed for efficient data reading and writing, while broom [@broom] was used for converting statistical objects into tidy data frames.
+Additional file management was facilitated by the here [@here] package.
+
+In order to build a model predicting the effect that various variables on the school-district-level have on their academic successes, we used three datasets from the Ontario government's Open Data portal: the School Board Achievements and Progress dataset [@achievements], the School Board Financial Reports dataset [@financial], and the School Information and Student Demographics dataset [@demographics], each recording different information on school district organised by district ID.
+
+After cleaning the data by selecting the desired variables from the three different datasets, combining them into one file, and removing N/A variables, we obtain data on the Four Year Graduation Rate, Total Expenses, Expenditure on Facilities, and Total Enrolment values for 50 different Ontarian school districts, as well as the percentage of students that are identified as low income, the percentage of students without a parent that has a school degree/certificate, the percentage of total expenses spent on facilities, and the expenses per quota in each of the aforementioned districts.
+
+As a measurement of academic success, the 4-year graduation rate was chosen due to its strong correlation with performance on the OSSLT, a standardized test used to assess Ontarian students' levels of academic literacy [@oasdi_graduation_requirements].
+
+## Measurement
+
+Data from the School Board Progress Report is either self-reported by schools or obtained by the EQAO (Education Quality and Accountability Office), as described by the Ministry of Education of Ontario [@ontario_education_faq].
+Schools are expected to self-report internal information such as class size changes and financial status, whereas indicators of academic progress on a district-wide-level, such as standardized testing results and graduation rates, fall under the authority of the EQAO.
+EQAO's data collection and processing methodology follows the Statistics Canada Quality Guidelines [@statcan_quality_assurance], and the aforementioned standardized assessments that EQAO records data on are administered by EQAO itself [@eqao_data_quality].
+
+Data regarding school districts' financial reports is collected by the Government of Ontario via the Education Financial Information System (EFIS).
+Using EFIS, school districts submit their financial information to the Ontario Ministry of Education, where they are then compiled into a single dataset.
+
+The School Information and Demographics dataset is based on a combination of school-submitted information and EQAO data.
+In light of privacy concerns such as data from this dataset potentially being traced back to individual students or small groups of students, random error and suppressing is utilised in this dataset [@ontario_sift_glossary].
+Specifically, when the number of students referred to by each individual cell in the table is less than 50, then the cell is marked as SP, or suppressed.
+Values concerning more students, on the other hand, have their percentages rounded up/down at random to a certain granularity, making sure that there is always a potential error bound of at most 5 students for each dataset.
+
+Several important limitations that must be acknowledged exist in the data collection and reporting processes for Ontario's education datasets.
+First, the self-reporting nature of some metrics by schools introduces potential inconsistencies in how different institutions interpret and report their data.
+Second, the privacy protection measures, while necessary, create inherent imprecision through data suppression and random rounding, particularly affecting analysis of smaller student populations or subgroups.
+The granularity of financial reporting through EFIS may also vary between school districts, potentially impacting the comparability of financial metrics.
+Additionally, the combination of multiple data sources (self-reported, EQAO, and demographic data) may lead to temporal misalignment, as different metrics might be collected at different times throughout the school year.
+Finally, the standardized testing results from EQAO, while following Statistics Canada guidelines, may not fully capture the diverse learning outcomes and educational experiences of students, particularly those from marginalized communities or with special educational needs.
+
+## Data Summary
+
+@tbl-summary presents an overview of the summary statistics of our key variables across Ontario school boards, and @tbl-snapshot presents a snapshot of the data itself.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-summary
+#| tbl-cap: "Summary Statistics of Key Variables"
 #| echo: false
+#| message: false
+#| warning: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+summary_stats %>%
+  mutate(
+    across(contains("Total Expenses"), ~./1e9),  # Convert to billions
+    across(contains("Total Enrolment"), ~./1000)  # Convert to thousands
+  ) %>%
+  pivot_longer(everything(), 
+               names_to = c("Variable", "Statistic"), 
+               names_pattern = "(.+)_(.+)") %>%
+  pivot_wider(names_from = Statistic, 
+             values_from = value) %>%
+  gt() %>%
+  fmt_number(
+    columns = c(Mean, SD, Min, Max, Median),
+    decimals = 2
+  ) %>%
+  cols_label(
+    Variable = "Variable",
+    Mean = "Mean",
+    SD = "Std. Dev.",
+    Min = "Minimum",
+    Max = "Maximum",
+    Median = "Median"
+  ) %>%
+  tab_footnote(
+    footnote = "Total Expenses shown in billions ($), Enrollment in thousands, percentages as raw values, expenses per student in dollars.",
+    locations = cells_column_labels(columns = Mean)
+  )
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+\FloatBarrier
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-snapshot
+#| tbl-cap: "Sample of First Five Rows from Ontario School Board Dataset"
 #| echo: false
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+# First half of columns
+knitr::kable(
+  analysis_data |>
+    select(`Board Number`,
+           `Board Name`,
+           Region,
+           City,
+           `Four Year Graduation Rate`,
+           `Total Expenses`) |>
+    slice(1:5) |>
+    mutate(Region = str_replace(Region, "Region ", "")),  # Remove the word "Region"
+  col.names = c("Board ID",
+                "Name",
+                "Region",
+                "City",
+                "Grad Rate",
+                "Total Expenses ($)"),
+  digits = c(0, 0, 0, 0, 3, 3),
+  booktabs = TRUE
+)
 
+# Second half of columns
+knitr::kable(
   analysis_data |>
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
+    select(`Expenditure on Facilities`,
+           `Total Enrolment`,
+           percent_low_income,
+           percent_no_degree,
+           percentage_spent_on_facilities,
+           expenses_per_quota) |>
+    slice(1:5),
+  col.names = c("Facility Expenses ($)",
+                "Enrollment",
+                "Low Income (%)",
+                "No Degree (%)",
+                "Facilities (%)",
+                "$/Student"),
+  digits = c(3, 0, 3, 3, 3, 3),
+  booktabs = TRUE
 )
@@
 
-# Discussion
+\FloatBarrier
 
-## First discussion point {#sec-first-point}
+The summary statistics reveal considerable variation across Ontario school boards.
+Four-year graduation rates average 82%, ranging from 59% to 96%.
+The diverse scales of the school boards, whether it be with respect to finances or population, are also evident: total expenses vary dramatically, from \$30 million to \$3.89 billion (with a mean of \$450 million), and so does total enrolment, which ranges from 710 students to 231,480 students, with a mean of 28,450 students.
+The percentage spent on facilities ranges from 6.06% to 12.55%, with a mean of 9.14%, while per-student expenses average \$18,730, ranging from \$13,524 to \$44,148.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Response Variable
 
-## Second discussion point
+### Four-Year Graduation Rate
 
-## Third discussion point
+The outcome variable is the four-year graduation rate for Ontario school boards, recorded as a percentage of students who graduate within four years of starting high school.
 
-## Weaknesses and next steps
+```{r}
+#| label: fig-grad
+#| fig-cap: Distribution of Four-Year Graduation Rates Across Ontario School Boards
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(analysis_data, aes(x = `Four Year Graduation Rate`)) +
+  geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
+  labs(x = "Four Year Graduation Rate (%)",
+       y = "Number of School Boards") +
+  theme_minimal()
+```
 
-Weaknesses and next steps should also be included.
+\FloatBarrier
+
+As displayed by @fig-grad, graduation rates across Ontario school boards follow an approximately normal distribution, with most boards achieving rates between 80% and 90%.
+The mean graduation rate is 82%, with rates ranging from 59% to 96%.
+
+## Predictor Variables
+
+The predictor variables in our analysis include the percentage of school board budget spent on facilities, per-student expenses, percentage of students without a parent holding a post-secondary degree, and percentage of students from low-income households.
+The distributions of these variables are as follows:
+
+### Facilities Spending
+
+```{r}
+#| label: fig-fac
+#| fig-cap: Distribution of facility spendings as a percentage of total spendings
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(analysis_data, aes(x = percentage_spent_on_facilities)) +
+  geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
+  labs(x = "Percentage Spent on Facilities",
+       y = "Number of School Boards") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+The percentage of budget allocated to facilities maintenance and operations ranges from 6.06% to 12.55%, with a mean of 9.14%.
+This metric captures the varying infrastructure needs and priorities across different school boards.
+
+### Per-Student Expenses
+
+```{r}
+#| label: fig-perquota
+#| fig-cap: Distribution of expenses per student
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(analysis_data, aes(x = expenses_per_quota)) +
+  geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
+  labs(x = "Expenses per Student ($)",
+       y = "Number of School Boards") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+As per @fig-perquota, per-student expenses show considerable variation across boards, averaging \$18,730 with a range from \$13,524 to \$44,148.
+This variation reflects differences in operational costs, programs offered, and resource allocation across different regions.
+
+### Parental Education
+
+```{r}
+#| label: fig-parents
+#| fig-cap: Percentage of students without parents holding a post-secondary degree
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(analysis_data, aes(x = percent_no_degree)) +
+  geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
+  labs(x = "Percentage of Students Without Parents Holding a Post-Secondary Degree",
+       y = "Number of School Boards") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+As can be seen in @fig-parents, the percentage of students whose parents do not hold a post-secondary degree or certificate varies from 1.85% to 10.66%, with a mean of 5.71%.
+This metric serves as an indicator of the educational background of the school board's community.
+
+### Low-Income Status
+
+```{r}
+#| label: fig-income
+#| fig-cap: Distribution of students in low-income families
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(analysis_data, aes(x = percent_low_income)) +
+  geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
+  labs(x = "Percentage of Students in Low Income Families",
+       y = "Number of School Boards") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+The percentage of students from low-income households ranges from 10.35% to 25.92%, with a mean of 15.77%.
+This variable provides insight into the socioeconomic composition of each school board's student population.
+
+### Enrolment
+
+```{r}
+#| label: fig-enroll
+#| fig-cap: Distribution of total enrolment across school boards
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(analysis_data, aes(x = `Total Enrolment`/1000)) +
+  geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
+  labs(x = "Total Enrolment (Thousands)",
+       y = "Number of School Boards") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+@fig-enroll shows that total enrolment across boards ranges from 710 to 231,480 students, with a mean of 28,450 students.
+This wide range reflects Ontario's diverse school board sizes, from small rural boards to large urban districts.
+
+## Feature Selection
+
+The selection of these specific predictor variables was guided by both theoretical relevance and statistical considerations regarding multicollinearity.
+Initial exploratory analysis revealed strong correlations between several potential predictors.
+For instance, total expenses and total enrollment showed an extremely high correlation (r > 0.95), as larger boards naturally have higher total expenses.
+Similarly, various socioeconomic indicators such as median household income, unemployment rates, and low-income status demonstrated substantial overlap in their variation.
+To address these multicollinearity concerns, we selected representative variables that captured distinct aspects of school board characteristics while minimizing redundancy:
+
+1. Per-student expenses was chosen over total expenses to control for board size and provide a standardized measure of financial resources.
+
+2.
+Percentage spent on facilities was retained as it represents a unique aspect of resource allocation independent of overall spending levels, as well as being reminiscent of the author's initial driving question of whether increased spending on libraries would correlate with better academic performance.
+
+3.
+Among socioeconomic indicators, the percentage of low-income students and parental education levels were selected as they captured different dimensions of socioeconomic status while maintaining relatively low correlation with each other (r < 0.4).
+
+4. Total enrolment was included as a control variable to account for potential scale effects that might influence graduation rates independently of other factors.
+
+Feature selection helps ensure our analysis avoids the statistical issues associated with multicollinearity while still capturing the key factors that may influence academic performance.
+
+## Note on Standardization
+
+While the data is presented in its raw form above, all variables are standardized (mean = 0, standard deviation = 1) for our subsequent modeling analysis to ensure comparability of coefficients and improve numerical stability, as well as to improve results obtained from the ridge and lasso regression models [@shrink_regression].
+This standardization helps interpret the relative importance of different predictors while maintaining their underlying relationships.
+
+# Model {#sec-model}
+
+## Model Set-up
+
+We evaluate several regression models to understand the relationship between school board characteristics and graduation rates:
+
+### Linear Regression
+Our baseline model is a simple linear regression:
+
+\begin{align} 
+y_i &= \beta_0 + \beta_1x_{1i} + \beta_2x_{2i} + \beta_3x_{3i} + \beta_4x_{4i} + \beta_5x_{5i} + \epsilon_i
+\end{align}
+
+Where $y_i$ is the Four Year Graduation Rate for school board $i$, $\epsilon_i$ is the error term, and $x_{1i}$ to $x_{5i}$ are the predictor variables: percentage of students without a parent holding a post-secondary degree, percentage of low-income students, percentage spent on facilities, expenses per student, and total enrollment.
+
+### Ridge Regression
+To address potential multicollinearity between our predictors, we implement Ridge regression:
+
+\begin{align}
+\min_{\beta} \sum_{i=1}^{n} (y_i - \beta_0 - \sum_{j=1}^{p} \beta_j x_{ij})^2 + \lambda \sum_{j=1}^{p} \beta_j^2
+\end{align}
+
+Where $\lambda$ is the regularization parameter chosen through cross-validation.
+
+### Lasso Regression
+We also implement Lasso regression for potential variable selection:
+
+\begin{align}
+\min_{\beta} \sum_{i=1}^{n} (y_i - \beta_0 - \sum_{j=1}^{p} \beta_j x_{ij})^2 + \lambda \sum_{j=1}^{p} |\beta_j|
+\end{align}
+
+### Polynomial Regression
+Our final and chosen model is a polynomial regression of degree 2:
+
+\begin{align}
+y_i &= \beta_0 + \sum_{j=1}^{5} (\beta_{j1}x_{ji} + \beta_{j2}x_{ji}^2) + \epsilon_i
+\end{align}
+
+Where $\beta_{j1}$ represents the linear term coefficient and $\beta_{j2}$ represents the quadratic term coefficient for each predictor.
+
+## Model Selection
+
+We chose the polynomial regression model for several reasons:
+
+1.
+**Non-linear Relationships**: Initial exploratory data analysis revealed non-linear relationships between several predictors and graduation rates, particularly for expenses per student and total enrollment.
+
+2. **Model Comparison**: Comparing Mean Squared Error (MSE) across models:
+
+[MSE SECTION HERE]
+
+The polynomial regression model showed the lowest MSE, indicating better fit to the data.
+
+3.
+**Interpretability**: While Ridge and Lasso regressions help with multicollinearity and variable selection respectively, they don't capture the non-linear relationships we observe in the data.
+The polynomial model allows us to interpret both linear and quadratic effects of our predictors on graduation rates.
+
+4.
+**Statistical Significance**: As shown in our model summary, several quadratic terms (particularly for expenses per quota) are statistically significant, confirming the value of including these higher-order terms.
+
+## Model Validation and Diagnostics
+
+### Residual Analysis
+```{r}
+#| label: fig-residuals
+#| fig-cap: Residual Diagnostics for Polynomial Regression Model
+#| echo: false
+#| message: false
+#| warning: false
+
+# Q-Q Plot
+ggplot(data.frame(residuals = residuals(poly_model)), aes(sample = residuals)) +
+  stat_qq() +
+  stat_qq_line() +
+  labs(title = "Normal Q-Q Plot",
+       x = "Theoretical Quantiles",
+       y = "Sample Quantiles") +
+  theme_minimal()
+
+# Residuals vs Fitted
+ggplot(data.frame(fitted = fitted(poly_model), 
+                 residuals = residuals(poly_model)), 
+       aes(x = fitted, y = residuals)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  labs(title = "Residuals vs Fitted Values",
+       x = "Fitted Values",
+       y = "Residuals") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+@fig-residuals shows our residual diagnostics.
+The Q-Q plot suggests that residuals are approximately normally distributed, while the Residuals vs Fitted plot shows no clear pattern, indicating homoscedasticity of residuals.
+
+### Model Performance Metrics
+
+```{r}
+#| label: tbl-metrics
+#| tbl-cap: Model Performance Metrics
+#| echo: false
+#| message: false
+#| warning: false
+
+metrics <- tibble(
+  Metric = c("R-squared", "RMSE"),
+  Value = c(summary(poly_model)$r.squared, 
+            sqrt(mean(residuals(poly_model)^2)))
+)
+
+metrics %>%
+  gt() %>%
+  fmt_number(columns = Value, decimals = 3)
+```
+
+\FloatBarrier
+
+As shown in @tbl-metrics, our model explains approximately 36.5% of the variance in graduation rates.
+While this R-squared value might seem modest, it is reasonable given the complexity of factors that influence educational outcomes.
+
+### Multicollinearity Check
+
+We examine Variance Inflation Factors (VIF) for the linear terms to assess multicollinearity:
+
+```{r}
+#| label: tbl-vif
+#| tbl-cap: Variance Inflation Factors for Linear Terms
+#| echo: false
+#| message: false
+#| warning: false
+
+# Calculate VIF for linear terms
+vif_values <- car::vif(lm_model)
+
+tibble(
+  Variable = names(vif_values),
+  VIF = vif_values
+) %>%
+  gt() %>%
+  fmt_number(columns = VIF, decimals = 2)
+```
+
+\FloatBarrier
+
+VIF values above 5 would indicate problematic multicollinearity. Our values suggest that multicollinearity is not a major concern in our model.
+
+### Cross-Validation
+
+To assess the model's predictive performance and guard against overfitting, we performed k-fold cross-validation:
+
+```{r}
+#| label: tbl-cv
+#| tbl-cap: Cross-Validation Results
+#| echo: false
+#| message: false
+#| warning: false
+
+set.seed(123)
+cv_results <- train(
+  `Four Year Graduation Rate` ~ .,
+  data = data_scaled,
+  method = "lm",
+  trControl = trainControl(method = "cv", number = 10)
+)
+
+tibble(
+  Metric = c("CV RMSE", "CV R-squared"),
+  Value = c(cv_results$results$RMSE, cv_results$results$Rsquared)
+) %>%
+  gt() %>%
+  fmt_number(columns = Value, decimals = 3)
+```
+
+\FloatBarrier
+
+The cross-validation results suggest that our model's performance is stable across different subsets of the data, with consistent RMSE values between training and testing sets.
+
+### Model Assumptions
+
+Our polynomial regression model relies on several key assumptions:
+
+1. **Linearity**: While we don't assume linear relationships between predictors and the response, we assume that the quadratic terms adequately capture the non-linear relationships.
+
+2. **Independence**: We assume that graduation rates of different school boards are independent of each other, which is reasonable given the administrative separation between boards.
+
+3. **Homoscedasticity**: As shown in our residual plot, the variance of residuals appears relatively constant across fitted values.
+
+4. **Normality**: The Q-Q plot suggests that residuals are approximately normally distributed.
+
+These diagnostics and validations support our choice of polynomial regression as the final model, though we acknowledge the limitations discussed in @sec-discussion.
+
+# Results {#sec-results}
+
+Our polynomial regression analysis reveals several significant relationships between school board characteristics and graduation rates. We present these results in order of statistical significance.
+
+## Primary Findings
+
+```{r}
+#| label: tbl-coefficients
+#| tbl-cap: Polynomial Regression Coefficients
+#| tbl-note: "All variables are standardized. Linear terms (1) represent direct relationships, while quadratic terms (2) represent non-linear effects."
+#| echo: false
+#| message: false
+#| warning: false
+
+tidy_poly <- broom::tidy(poly_model) %>%
+  select(-statistic) %>%
+  mutate(
+    term = str_remove(term, "poly\\("),
+    term = str_remove(term, ", 2\\)"),
+    term = str_remove(term, "\\)")
+  )
+
+tidy_poly %>%
+  gt() %>%
+  fmt_number(columns = c(estimate, std.error, p.value), decimals = 3) %>%
+  cols_label(
+    term = "Variable",
+    estimate = "Estimate",
+    std.error = "Std. Error",
+    p.value = "P-value"
+  )
+```
+
+\FloatBarrier
+
+### Parental Education
+
+The percentage of households without post-secondary degrees shows the strongest relationship with graduation rates (β = -3.627, p = 0.003).
+This linear relationship suggests that for every standard deviation increase in the percentage of households without degrees, graduation rates decrease by approximately 3.6 percentage points, holding other factors constant.
+
+### Facilities Spending
+
+```{r}
+#| label: fig-facilities
+#| fig-cap: Relationship Between Facilities Spending and Graduation Rates
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(data_scaled, aes(x = percentage_spent_on_facilities, 
+                        y = `Four Year Graduation Rate`)) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "lm", 
+              formula = y ~ poly(x, 2), 
+              se = TRUE) +
+  labs(x = "Percentage Spent on Facilities (Standardized)",
+       y = "Graduation Rate (Standardized)") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+The percentage spent on facilities shows a significant negative linear relationship (β = -2.402, p = 0.026), as illustrated in @fig-facilities.
+This suggests that boards allocating a larger proportion of their budget to facilities tend to have lower graduation rates, possibly indicating trade-offs between infrastructure spending and other educational resources.
+
+### Per-Student Expenses
+
+```{r}
+#| label: fig-expenses
+#| fig-cap: Non-linear Relationship Between Per-Student Expenses and Graduation Rates
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(data_scaled, aes(x = expenses_per_quota, 
+                        y = `Four Year Graduation Rate`)) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "lm", 
+              formula = y ~ poly(x, 2), 
+              se = TRUE) +
+  labs(x = "Expenses per Student (Standardized)",
+       y = "Graduation Rate (Standardized)") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+Per-student expenses show a significant quadratic relationship (β = 3.444, p = 0.013), as shown in @fig-expenses.
+The U-shaped relationship suggests that while initial increases in per-student spending may be associated with lower graduation rates, there appears to be a threshold beyond which additional spending becomes positively associated with graduation rates.
+
+## Secondary Findings
+
+### Total Expenses and Enrollment
+
+While neither total expenses nor total enrollment showed statistically significant relationships with graduation rates at the conventional p < 0.05 level, both variables exhibited potentially meaningful trends:
+
+```{r}
+#| label: fig-size
+#| fig-cap: Relationships Between School Board Size Metrics and Graduation Rates
+#| layout-ncol: 2
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(data_scaled, aes(x = `Total Expenses`, 
+                        y = `Four Year Graduation Rate`)) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "lm", 
+              formula = y ~ poly(x, 2), 
+              se = TRUE) +
+  labs(x = "Total Expenses (Standardized)",
+       y = "Graduation Rate (Standardized)") +
+  theme_minimal()
+
+ggplot(data_scaled, aes(x = `Total Enrolment`, 
+                        y = `Four Year Graduation Rate`)) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "lm", 
+              formula = y ~ poly(x, 2), 
+              se = TRUE) +
+  labs(x = "Total Enrollment (Standardized)",
+       y = "Graduation Rate (Standardized)") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+As shown in @fig-size, both total expenses and enrollment suggest potential non-linear relationships with graduation rates, though these relationships did not reach statistical significance in our model.
+
+### Model Fit
+
+```{r}
+#| label: fig-fit
+#| fig-cap: Predicted vs Actual Graduation Rates
+#| echo: false
+#| message: false
+#| warning: false
+ggplot(data_scaled, aes(x = fitted(poly_model), 
+                        y = `Four Year Graduation Rate`)) +
+  geom_point(alpha = 0.6) +
+  geom_abline(intercept = 0, slope = 1, 
+              color = "red", linetype = "dashed") +
+  labs(x = "Predicted Graduation Rate (Standardized)",
+       y = "Actual Graduation Rate (Standardized)") +
+  theme_minimal()
+```
+
+\FloatBarrier
+
+@fig-fit shows the relationship between predicted and actual graduation rates.
+The model explains approximately 40.4% of the variance in graduation rates (R² = 0.404), suggesting that while our identified factors are important, there remain other unobserved variables that influence graduation rates.
+
+# Discussion {#sec-discussion}
+
+## Key Findings and Implications
+
+Our analysis reveals several important insights about the factors influencing graduation rates in Ontario school boards.
+The most striking finding is the strong negative relationship between the percentage of households without post-secondary degrees and graduation rates.
+This relationship suggests an intergenerational aspect to educational achievement, where students from communities with lower educational attainment face additional challenges in completing their secondary education.
+
+The complex relationship between per-student expenses and graduation rates is particularly noteworthy.
+The U-shaped relationship we observed suggests that simply increasing funding may not always lead to better outcomes.
+This finding aligns with previous research suggesting that the effectiveness of educational spending depends heavily on how resources are allocated [@educationspending].
+School boards might benefit from examining the spending patterns of high-performing boards that operate in the "efficient" region of this curve.
+
+## Resource Allocation Trade-offs
+
+The negative relationship between facilities spending and graduation rates raises important questions about resource allocation.
+While maintaining adequate facilities is crucial for educational delivery, our findings suggest that boards allocating a larger proportion of their budget to facilities tend to have lower graduation rates.
+This could indicate that:
+
+1. Some boards might be forced to prioritize urgent infrastructure needs over other educational resources
+2. Older facilities requiring more maintenance might be concentrated in areas facing other socioeconomic challenges
+3. There might be an optimal balance between infrastructure investment and direct educational spending
+
+## Limitations
+
+Several limitations of our analysis should be noted:
+
+### Data Constraints
+1. Our analysis is cross-sectional, looking at a single academic year (2019-2020)
+2. The COVID-19 pandemic may have influenced both spending patterns and graduation rates during this period
+3. We lack data on important factors such as:
+   - Teacher qualifications and experience
+   - Specific program offerings
+   - Student-level characteristics
+
+### Methodological Considerations
+1. Our polynomial regression model, while capturing non-linear relationships, may not fully represent more complex interactions between variables
+2. The R-squared value of 0.404 suggests that substantial variation in graduation rates remains unexplained by our model
+3. The standardization of variables, while necessary for our analysis, makes direct interpretation of effect sizes less intuitive
+
+## Future Research Directions
+
+Several promising avenues for future research emerge from our findings:
+
+1. **Longitudinal Analysis**: Examining how these relationships evolve over time could provide insights into the long-term effects of different spending patterns and policy changes.
+
+2.
+**Program-Level Analysis**: Investigating specific educational programs and interventions could help identify which approaches are most effective at improving graduation rates, particularly in communities with lower educational attainment.
+
+3.
+**Resource Allocation Optimization**: Developing models to help school boards optimize their resource allocation decisions, particularly regarding the balance between facilities maintenance and other educational spending.
+
+4.
+**Policy Implications**: Studying how provincial funding formulas might be adjusted to better support boards facing particular challenges, such as aging infrastructure or high concentrations of socioeconomic disadvantage.
+
+## Broader Context
+
+Our findings contribute to the ongoing discussion about educational equity in Ontario.
+The strong relationship between community educational attainment and graduation rates suggests that breaking cycles of educational disadvantage may require interventions that extend beyond the school system itself.
+This might include:
+
+1. Community-based programs to support adult education
+2. Enhanced early childhood education initiatives
+3. Targeted support for communities with historically lower educational attainment
+4. Integrated approaches that address both educational and broader socioeconomic factors
+
+The complex relationships we've identified between financial resources and educational outcomes also suggest that policy makers should consider more nuanced approaches to school funding, moving beyond simple per-student funding formulas to consider the specific contexts and challenges faced by different school boards.
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage
@@
 https://data.ontario.ca/dataset/school-information-and-student-demographics/resource/e0e90bd5-d662-401a-a6d2-60d69ac89d14
-
-https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1810000501
-
-https://stats.stackexchange.com/questions/11676/pseudo-r-squared-formula-for-glms

### 2024-12-04 scripts/00-simulate_data.R

@@ -30,9 +30,10 @@ simulated_data <- tibble(
   percent_low_income = round(runif(n_rows, 10, 50), 3),                            # Low-income percentage between 10% and 50%
   percent_no_degree = round(runif(n_rows, 5, 30), 3)                               # No-degree percentage between 5% and 30%
 ) %>%
-  # Calculate percentage_spent_on_facilities based on Total Expenses and Expenditure on Facilities
+  # Calculate additional columns
   mutate(
-    percentage_spent_on_facilities = round((`Expenditure on Facilities` / `Total Expenses`) * 100, 3)
+    percentage_spent_on_facilities = round((`Expenditure on Facilities` / `Total Expenses`) * 100, 3),
+    expenses_per_quota = round(`Total Expenses` / `Total Enrolment`, 3)             # Calculate expenses per quota
   )
 
 #### Save simulated data ####
@@ -42,5 +43,3 @@ write_parquet(simulated_data, here::here("./data/simulated_data/simulated_data.p
 #### Summary ####
 # Print a message indicating the data simulation is complete
 cat("Simulated data saved to './data/simulated_data/simulated_data.parquet'.\n")
-
-

### 2024-12-04 scripts/01-test_simulated_data.R

@@ -14,6 +14,12 @@ library(testthat)
 # Load simulated data from the parquet file
 simulated_data <- read_parquet(here::here("./data/simulated_data/simulated_data.parquet"))
 
+#### Add `expenses_per_quota` column ####
+simulated_data <- simulated_data %>%
+  mutate(
+    expenses_per_quota = round(`Total Expenses` / `Total Enrolment`, 3)  # Add and round to 3 decimals
+  )
+
 #### Define sanity check function ####
 # Function to perform tests on the dataset
 run_sanity_checks <- function(data, dataset_name) {
@@ -28,24 +34,28 @@ run_sanity_checks <- function(data, dataset_name) {
                 info = paste(dataset_name, "does not have between 5 and 1000 rows")
     )
     
-    # Ensure the dataset has exactly 11 columns
-    expect_equal(ncol(data), 11,
-                 info = paste(dataset_name, "does not have exactly 11 columns")
+    # Ensure the dataset has exactly 12 columns (new column added)
+    expect_equal(ncol(data), 12,
+                 info = paste(dataset_name, "does not have exactly 12 columns")
     )
     
-    # Confirm all required columns are present
+    # Confirm all required columns are present, including the new column
     required_cols <- c(
       "Board Number", "Board Name", "Region", "City",
       "Four Year Graduation Rate", "Total Expenses",
       "Expenditure on Facilities", "Total Enrolment",
-      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities"
+      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities", 
+      "expenses_per_quota"
     )
     expect_true(all(required_cols %in% colnames(data)),
                 info = paste(dataset_name, "is missing required columns")
     )
     
     #### Data type tests ####
-    # Check the types of key variables
+    # Check the types of key variables, including the new column
+    expect_type(data$expenses_per_quota, "double")
+    
+    # Existing type checks
     expect_type(data$`Board Number`, "character")
     expect_type(data$`Board Name`, "character")
     expect_type(data$Region, "character")
@@ -59,26 +69,25 @@ run_sanity_checks <- function(data, dataset_name) {
     expect_type(data$percentage_spent_on_facilities, "double")
     
     #### Value range tests ####
-    # Verify graduation rates are between 0 and 1
+    # Ensure expenses_per_quota is non-negative
+    expect_true(all(data$expenses_per_quota >= 0),
[…]

### 2024-12-04 scripts/03-clean_data.R

@@ -89,19 +89,21 @@ final_data <- final_data%>%
 final_data <- final_data%>%
   rename("Four Year Graduation Rate" = `Four Year Graduation Rate 2019-2020 Grade 9 Cohort`)
 
-# Make sure numeric variables are numeric, and add a new column
+# Make sure numeric variables are numeric, and add new columns
 final_data <- final_data %>%
   mutate(
     `Expenditure on Facilities` = as.numeric(`Expenditure on Facilities`),
     `Four Year Graduation Rate` = as.numeric(`Four Year Graduation Rate`),
     `Total Expenses` = as.numeric(`Total Expenses`),
-    percentage_spent_on_facilities = (`Expenditure on Facilities` / `Total Expenses`) * 100
+    percentage_spent_on_facilities = (`Expenditure on Facilities` / `Total Expenses`) * 100,
+    expenses_per_quota = `Total Expenses` / `Total Enrolment` # Add expenses_per_quota
   )
 
 # Round everything to 3 decimal places
 final_data <- final_data %>%
   mutate(across(where(is.numeric), ~ round(. , 3)))
 
+# Remove rows with NA values
 final_data <- na.omit(final_data)
 
 # Save data as a parquet

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -28,9 +28,9 @@ run_sanity_checks <- function(data, dataset_name) {
                 info = paste(dataset_name, "does not have between 5 and 1000 rows")
     )
     
-    # Ensure the dataset has exactly 11 columns
-    expect_equal(ncol(data), 11,
-                 info = paste(dataset_name, "does not have exactly 11 columns")
+    # Ensure the dataset has exactly 12 columns
+    expect_equal(ncol(data), 12,
+                 info = paste(dataset_name, "does not have exactly 12 columns")
     )
     
     # Confirm all required columns are present
@@ -38,7 +38,8 @@ run_sanity_checks <- function(data, dataset_name) {
       "Board Number", "Board Name", "Region", "City",
       "Four Year Graduation Rate", "Total Expenses",
       "Expenditure on Facilities", "Total Enrolment",
-      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities"
+      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities",
+      "expenses_per_quota"
     )
     expect_true(all(required_cols %in% colnames(data)),
                 info = paste(dataset_name, "is missing required columns")
@@ -57,6 +58,7 @@ run_sanity_checks <- function(data, dataset_name) {
     expect_type(data$percent_low_income, "double")
     expect_type(data$percent_no_degree, "double")
     expect_type(data$percentage_spent_on_facilities, "double")
+    expect_type(data$expenses_per_quota, "double")
     
     #### Value range tests ####
     # Verify graduation rates are between 0 and 1
@@ -78,6 +80,12 @@ run_sanity_checks <- function(data, dataset_name) {
                 info = paste(dataset_name, "`percentage_spent_on_facilities` is not calculated correctly")
     )
     
+    # Ensure `expenses_per_quota` is calculated correctly
+    calculated_quota <- data$`Total Expenses` / data$`Total Enrolment`
+    expect_true(all(abs(calculated_quota - data$expenses_per_quota) < 1e-3),
+                info = paste(dataset_name, "`expenses_per_quota` is not calculated correctly")
+    )
+    
     # Confirm income and education percentages are between 0 and 100
     expect_true(all(data$percent_low_income >= 0 & data$percent_low_income <= 100),
                 info = paste(dataset_name, "`percent_low_income` values are out of range")
@@ -95,7 +103,7 @@ run_sanity_checks <- function(data, dataset_name) {
[…]

### 2024-12-04 scripts/05-model.R

@@ -8,106 +8,82 @@
 #### Workspace setup ####
 library(tidyverse)
 library(arrow)
-library(brms)
 library(broom)
 library(caret)
+library(glmnet)      # For Ridge and Lasso regression
 
 #### Load data ####
-# Read analysis data from the parquet file
 analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.parquet"))
-
-# Rename columns to remove spaces
 colnames(analysis_data) <- gsub(" ", "_", colnames(analysis_data))
 
-# Check the new column names
-colnames(analysis_data)
-
 #### Data Preprocessing ####
-# Select relevant columns for modeling
 data <- analysis_data %>%
   dplyr::select(Four_Year_Graduation_Rate, 
-         Total_Expenses, 
-         percent_no_degree, 
-         percent_low_income, 
-         percentage_spent_on_facilities)
+                percent_no_degree, 
+                percent_low_income, 
+                percentage_spent_on_facilities,
+                expenses_per_quota,
+                Total_Enrolment)
 
-# Normalize data using caret package for scaling
+# Normalize data
 data_scaled <- data %>%
   mutate(across(everything(), ~ scale(.) %>% as.vector()))
 
+# Prepare data for glmnet
+x <- model.matrix(Four_Year_Graduation_Rate ~ ., data_scaled)[, -1]
+y <- data_scaled$Four_Year_Graduation_Rate
+
 #### Linear Regression Model ####
-# Fit the linear regression model
-lm_model <- lm(Four_Year_Graduation_Rate ~ Total_Expenses + percent_no_degree + percent_low_income + percentage_spent_on_facilities, data = data_scaled)
+lm_model <- lm(Four_Year_Graduation_Rate ~ ., data = data_scaled)
 
-# Print model summary
-lm_summary <- summary(lm_model)
-lm_stats <- broom::tidy(lm_model)
+#### Ridge Regression ####
+ridge_cv <- cv.glmnet(x, y, alpha = 0)
+ridge_best_lambda <- ridge_cv$lambda.min
+ridge_model <- glmnet(x, y, alpha = 0, lambda = ridge_best_lambda)
 
-# Print R-squared and p-values
-print("Linear Model Summary:")
-print(lm_summary)
+#### Lasso Regression ####
+lasso_cv <- cv.glmnet(x, y, alpha = 1)
+lasso_best_lambda <- lasso_cv$lambda.min
[…]

### 2024-12-04 paper/paper.qmd
updated miscellaneous aspects of the project like sketches and README

@@
 ---
-title: "Facilities Spending and Socioeconomic Factors Show Strong Correlation with Ontario School District Graduation Rates"
+title: "A 1% Increase in Spending on Facilities Decreases Graduation Rates by 0.2%: An Analysis on Ontario School Districts and Related Predictors"
 author: Andrew Goh
@@
 abstract: "Using data from Ontario's school board achievement, financial, and demographic datasets, we analyze the relationship between various district-level factors and four-year graduation rates.
-Through polynomial regression analysis, we find that the percentage of households without post-secondary degrees and facilities spending have significant negative correlations with graduation rates, while per-student expenses show a non-linear relationship.
-These findings suggest that socioeconomic factors and resource allocation decisions significantly impact academic outcomes, with implications for education policy and resource distribution in Ontario school districts."
+Through polynomial regression analysis, we find that the percentage of households without post-secondary degrees and relative spending on facilities have significant negative correlations with graduation rates (-0.136, -0.197), while per-student expenses show a non-linear relationship.
+These findings suggest that socioeconomic factors and resource allocation decisions have statistically significant impacts on academic outcomes, with implications for education policy and resource distribution in Ontario school districts."
 format: pdf
@@
 lm_model <- readRDS(here("models", "lm_model.rds"))
+ridge_model <- readRDS(here("models", "ridge_model.rds"))
+lasso_model <- readRDS(here("models", "lasso_model.rds"))
 
@@
 The relationship between educational resources and academic outcomes has been a central focus of education policy research, particularly in publicly funded school systems.
-In Ontario, where school boards receive varying levels of funding and serve communities with diverse socioeconomic profiles, understanding how different factors influence student success is crucial for effective policy-making [@ontarioeducation].
+In Ontario, where school boards receive varying levels of funding and serve communities with diverse socioeconomic profiles, understanding how different factors influence student success is crucial for effective policy-making [@campbell].
 
-Recent studies have suggested that both financial resources and community characteristics play significant roles in determining student outcomes [@educationresources; @socioeconomic].
+Recent studies have suggested that both financial resources and community characteristics play significant roles in determining student outcomes [@financial_support_academic_performance; @socioeconomic_status_achievement].
 However, the specific mechanisms through which these factors interact, particularly in the Ontario context, remain unclear.
-This gap in understanding is particularly relevant given Ontario's significant investment in education, with annual spending exceeding $31 billion [@ontariobudget].
+This gap in understanding is particularly relevant given Ontario's significant investment in education, with annual spending projected to be approximately \$37.6 billion [@opsba_budget_overview].
 
 Our paper examines how various district-level factors—including financial allocations, socioeconomic indicators, and operational characteristics—correlate with four-year graduation rates across Ontario school boards.
-We focus specifically on the relationships between graduation rates and six key variables: total expenses, percentage of households without post-secondary degrees, percentage of low-income households, percentage of budget spent on facilities, per-student expenses, and total enrollment.
+We focus specifically on the relationships between graduation rates and five key variables: the percentage of households without post-secondary degrees, the percentage of low-income households, the percentage of the total budget spent on facilities, per-student expenses, and total enrolment.
 
@@
 
-The results suggest that socioeconomic factors, particularly the percentage of households without post-secondary degrees, have a significant negative linear relationship with graduation rates.
-Additionally, we find that the relationship between per-student expenses and graduation rates follows a U-shaped curve, suggesting diminishing returns up to a certain point.
-These findings have important implications for how educational resources might be most effectively allocated to improve student outcomes.
+The results suggest that socioeconomic factors, particularly the percentage of households without post-secondary degrees, show the strongest relationship with graduation rates, with each percentage point increase associated with a 0.136 percentage point decrease in graduation rates.
+We also find that the percentage of budget spent on facilities has a significant negative relationship with graduation rates, with each additional percentage point allocated to facilities associated with a 0.197 percentage point decrease in graduation rates.
+While per-student expenses show a U-shaped relationship with graduation rates, this relationship did not reach statistical significance at conventional levels.
+Our model explains approximately 36.5% of the variance in graduation rates, indicating that while these factors are important, there remain other unobserved variables influencing student outcomes.
 
@@
 ## Overview
+
 We use the statistical programming language R [@citer] for the analysis and presentation of the project.
@@
 
-@tbl-summary presents an overview of the summary statistics of our key variables across Ontario school boards, and @tbl-snapshot presents a snapshot of the data itself.
+@tbl-summary presents an overview of the summary statistics of our key variables across Ontario school boards, and @tbl-snapshot1 and @tbl-snapshot2 presents a snapshot of the data itself.
 
@@
   tab_footnote(
-    footnote = "Total Expenses shown in billions ($), Enrollment in thousands, percentages as raw values, expenses per student in dollars.",
+    footnote = "Total Expenses shown in billions ($), Enrolment in thousands, percentages as raw values, expenses per student in dollars.",
     locations = cells_column_labels(columns = Mean)
@@
 ```{r}
-#| label: tbl-snapshot
-#| tbl-cap: "Sample of First Five Rows from Ontario School Board Dataset"
+#| label: tbl-snapshot1
+#| tbl-cap: "Sample of First Five Rows from Ontario School Board Dataset (Part 1)"
 #| echo: false
@@
     slice(1:5) |>
-    mutate(Region = str_replace(Region, "Region ", "")),  # Remove the word "Region"
+    mutate(Region = str_replace(Region, "Region ", "")),
   col.names = c("Board ID",
@@
   digits = c(0, 0, 0, 0, 3, 3),
-  booktabs = TRUE
-)
+  booktabs = TRUE)
+```
+```{r}
+#| label: tbl-snapshot2
+#| tbl-cap: "Sample of First Five Rows from Ontario School Board Dataset (Part 2)"
 
@@
   col.names = c("Facility Expenses ($)",
-                "Enrollment",
+                "Enrolment",
                 "Low Income (%)",
@@
   digits = c(3, 0, 3, 3, 3, 3),
-  booktabs = TRUE
-)
+  booktabs = TRUE)
 ```
@@
 The percentage of students from low-income households ranges from 10.35% to 25.92%, with a mean of 15.77%.
-This variable provides insight into the socioeconomic composition of each school board's student population.
+This variable describes the socioeconomic composition of each school board's student population, and we can see from @fig-income the diversity of socioeconomic compositions *between* school districts.
 
@@
 Initial exploratory analysis revealed strong correlations between several potential predictors.
-For instance, total expenses and total enrollment showed an extremely high correlation (r > 0.95), as larger boards naturally have higher total expenses.
+For instance, total expenses and total enrolment showed an extremely high correlation (r \> 0.95), as larger boards naturally have higher total expenses.
 Similarly, various socioeconomic indicators such as median household income, unemployment rates, and low-income status demonstrated substantial overlap in their variation.
@@
 3.
-Among socioeconomic indicators, the percentage of low-income students and parental education levels were selected as they captured different dimensions of socioeconomic status while maintaining relatively low correlation with each other (r < 0.4).
+Among socioeconomic indicators, the percentage of low-income students and parental education levels were selected as they captured different dimensions of socioeconomic status while maintaining relatively low correlation with each other (r \< 0.4).
 
@@
 
-We evaluate several regression models to understand the relationship between school board characteristics and graduation rates:
+We evaluate the performance of several regression models to understand the relationship between school board characteristics and graduation rates:
 
 ### Linear Regression
-Our baseline model is a simple linear regression:
 
+We begin with our baseline model, a simple linear regression model:
+
+```{=tex}
 \begin{align} 
@@
 \end{align}
+```
+Where $y_i$ is the Four Year Graduation Rate for school board $i$, $\epsilon_i$ is the error term, and $x_{1i}$ to $x_{5i}$ are the predictor variables: percentage of students without a parent holding a post-secondary degree, percentage of low-income students, percentage spent on facilities, expenses per student, and total enrolment.
 
-Where $y_i$ is the Four Year Graduation Rate for school board $i$, $\epsilon_i$ is the error term, and $x_{1i}$ to $x_{5i}$ are the predictor variables: percentage of students without a parent holding a post-secondary degree, percentage of low-income students, percentage spent on facilities, expenses per student, and total enrollment.
+### Shrinkage Regression
 
-### Ridge Regression
-To address potential multicollinearity between our predictors, we implement Ridge regression:
+To address potential multicollinearity and model overcomplexity between our predictors, we also implemented Ridge and Lasso regression models, respectively:
 
+```{=tex}
 \begin{align}
@@
 \end{align}
-
-Where $\lambda$ is the regularization parameter chosen through cross-validation.
-
-### Lasso Regression
-We also implement Lasso regression for potential variable selection:
-
+```
+```{=tex}
 \begin{align}
@@
 \end{align}
+```
+Where $\lambda$ is the regularization parameter chosen through cross-validation and the x-values are the same as above.
 
 ### Polynomial Regression
-Our final and chosen model is a polynomial regression of degree 2:
 
+Finally, a polynomial regression model of degree 2 was fitted to the data to account for non-linear relationships between the data. The model is structured as follows:
+
+```{=tex}
 \begin{align}
@@
 \end{align}
-
+```
 Where $\beta_{j1}$ represents the linear term coefficient and $\beta_{j2}$ represents the quadratic term coefficient for each predictor.
@@
 
-We chose the polynomial regression model for several reasons:
+After comparing the results of the models and assessing how well they each fit to the data respectively, the polynomial regression model was chosen due to multiple reasons:
 
 1.
-**Non-linear Relationships**: Initial exploratory data analysis revealed non-linear relationships between several predictors and graduation rates, particularly for expenses per student and total enrollment.
+**Non-linear Relationships**: Initial exploratory data analysis revealed non-linear relationships between several predictors and graduation rates, particularly for expenses per student and total enrolment.
+This suggests that a polynomial regression model potentially can capture more trends in the data, more accurately, compared to a linear model, though the latter is simpler.
+The tradeoff in this case is sensible due to the added model complexity actively capturing statistically significant trends in the data.
 
-2. **Model Comparison**: Comparing Mean Squared Error (MSE) across models:
+2.  **Model Comparison**: Comparing Root Mean Squared Error (RMSE) across models:
 
-[MSE SECTION HERE]
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+library(broom)
+library(caret)
+library(glmnet)      # For Ridge and Lasso regression
 
-The polynomial regression model showed the lowest MSE, indicating better fit to the data.
+#### Load data ####
+analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.parquet"))
+
+#### Data Preprocessing ####
+data <- analysis_data %>%
+  dplyr::select(`Four Year Graduation Rate`, 
+                percent_no_degree, 
+                percent_low_income, 
+                percentage_spent_on_facilities,
+                expenses_per_quota,
+                `Total Enrolment`)
+
+# Normalize data
+data_scaled <- data %>%
+  mutate(across(everything(), ~ scale(.) %>% as.vector()))
+
+# Prepare data for glmnet
+x <- model.matrix(`Four Year Graduation Rate` ~ ., data_scaled)[, -1]
+y <- data_scaled$`Four Year Graduation Rate`
+
+#### Linear Regression Model ####
+lm_model <- lm(`Four Year Graduation Rate` ~ ., data = data_scaled)
+
+#### Ridge Regression ####
+ridge_cv <- cv.glmnet(x, y, alpha = 0)
+ridge_best_lambda <- ridge_cv$lambda.min
+ridge_model <- glmnet(x, y, alpha = 0, lambda = ridge_best_lambda)
+
+#### Lasso Regression ####
+lasso_cv <- cv.glmnet(x, y, alpha = 1)
+lasso_best_lambda <- lasso_cv$lambda.min
+lasso_model <- glmnet(x, y, alpha = 1, lambda = lasso_best_lambda)
+
+#### Polynomial Regression ####
+poly_model <- lm(`Four Year Graduation Rate` ~ 
+                   poly(percent_no_degree, 2) + 
+                   poly(percent_low_income, 2) +
+                   poly(percentage_spent_on_facilities, 2) +
+                   poly(expenses_per_quota, 2) +
+                   poly(`Total Enrolment`, 2),
+                 data = data_scaled)
+
+#### Model Comparison with MSE ####
+# Predictions
+lm_preds <- predict(lm_model, newdata = data_scaled)
+poly_preds <- predict(poly_model, newdata = data_scaled)
+ridge_preds <- predict(ridge_model, newx = x, s = ridge_best_lambda)
+lasso_preds <- predict(lasso_model, newx = x, s = lasso_best_lambda)
+
+# Calculate RMSE
+lm_rmse <- sqrt(mean((lm_preds - data_scaled$`Four Year Graduation Rate`)^2))
+poly_rmse <- sqrt(mean((poly_preds - data_scaled$`Four Year Graduation Rate`)^2))
+ridge_rmse <- sqrt(mean((ridge_preds - y)^2))
+lasso_rmse <- sqrt(mean((lasso_preds - y)^2))
+
+# Model Comparison Table
+model_comparison <- tibble(
+  Model = c("Linear Regression", "Polynomial Regression", "Ridge", "Lasso"),
+  Metric = c(
+    paste("RMSE: ", round(lm_rmse, 3)),
+    paste("RMSE: ", round(poly_rmse, 3)),
+    paste("RMSE: ", round(ridge_rmse, 3)),
+    paste("RMSE: ", round(lasso_rmse, 3))
+  )
+)
+
+print(model_comparison)
+```
+
+The polynomial regression model showed the lowest MSE, which indicates that it has a better fit to the data and is hence a "better" model to use in this case [@validation].
 
@@
 ### Residual Analysis
+
 ```{r}
-#| label: fig-residuals
-#| fig-cap: Residual Diagnostics for Polynomial Regression Model
+#| label: fig-qq
+#| fig-cap: Normal QQ Plot for Polynomial Regression Model
 #| echo: false
@@
   theme_minimal()
+```
+```{r}
+#| label: fig-residuals
+#| fig-cap: Residuals vs Fitted Values Plot for Polynomial Regression Model
+#| echo: false
+#| message: false
+#| warning: false
 
@@
 
-@fig-residuals shows our residual diagnostics.
+@fig-qq and @fig-residuals show our residual diagnostics.
 The Q-Q plot suggests that residuals are approximately normally distributed, while the Residuals vs Fitted plot shows no clear pattern, indicating homoscedasticity of residuals.
@@
 
-Our polynomial regression analysis reveals several significant relationships between school board characteristics and graduation rates. We present these results in order of statistical significance.
+We present the results from our polynomial regression analysis in order of statistical significance.
 
@@
 The percentage of households without post-secondary degrees shows the strongest relationship with graduation rates (β = -3.627, p = 0.003).
-This linear relationship suggests that for every standard deviation increase in the percentage of households without degrees, graduation rates decrease by approximately 3.6 percentage points, holding other factors constant.
+After converting from standardized units back to real-world units, this relationship suggests that for every one percentage point increase in households without post-secondary degrees, graduation rates decrease by approximately 0.136 percentage points, holding other factors constant.
 
@@
 
-The percentage spent on facilities shows a significant negative linear relationship (β = -2.402, p = 0.026), as illustrated in @fig-facilities.
-This suggests that boards allocating a larger proportion of their budget to facilities tend to have lower graduation rates, possibly indicating trade-offs between infrastructure spending and other educational resources.
+The percentage spent on facilities shows a significant negative linear relationship (β = -2.435, p = 0.023), as illustrated in @fig-facilities.
+After converting from standardized units back to real-world units, this suggests that for each additional percentage point of budget allocated to facilities, graduation rates decrease by about 0.197 percentage points, holding other factors constant.
+In other words, boards allocating a larger proportion of their budget to facilities tend to have lower graduation rates, possibly indicating trade-offs between infrastructure spending and other educational resources.
 
@@
 
-Per-student expenses show a significant quadratic relationship (β = 3.444, p = 0.013), as shown in @fig-expenses.
-The U-shaped relationship suggests that while initial increases in per-student spending may be associated with lower graduation rates, there appears to be a threshold beyond which additional spending becomes positively associated with graduation rates.
+Per-student expenses show a marginally significant quadratic relationship (β = 1.932, p = 0.085), as shown in @fig-expenses.
+After converting from standardized units back to real-world units, while not significant at the conventional p < 0.05 level, there appears to be a U-shaped trend in the relationship between per-student spending and graduation rates.
+This suggests that graduation rates tend to decrease with initial increases in per-student spending up to a certain threshold, after which additional spending is associated with increasing graduation rates.
+However, given the lack of statistical significance, these trends should be interpreted with caution.
 
@@
 
-### Total Expenses and Enrollment
-
-While neither total expenses nor total enrollment showed statistically significant relationships with graduation rates at the conventional p < 0.05 level, both variables exhibited potentially meaningful trends:
+### Total Expenses and Enrolment
 
@@
 #| fig-cap: Relationships Between School Board Size Metrics and Graduation Rates
-#| layout-ncol: 2
 #| echo: false
@@
 #| warning: false
-ggplot(data_scaled, aes(x = `Total Expenses`, 
-                        y = `Four Year Graduation Rate`)) +
-  geom_point(alpha = 0.6) +
-  geom_smooth(method = "lm", 
-              formula = y ~ poly(x, 2), 
-              se = TRUE) +
-  labs(x = "Total Expenses (Standardized)",
-       y = "Graduation Rate (Standardized)") +
-  theme_minimal()
-
 ggplot(data_scaled, aes(x = `Total Enrolment`, 
@@
               se = TRUE) +
-  labs(x = "Total Enrollment (Standardized)",
+  labs(x = "Total Enrolment (Standardized)",
        y = "Graduation Rate (Standardized)") +
@@
 
-As shown in @fig-size, both total expenses and enrollment suggest potential non-linear relationships with graduation rates, though these relationships did not reach statistical significance in our model.
+As shown in @fig-size, Total Enrolment showed no statistically significant relationship with graduation rates (linear term: β = 0.249, p = 0.842; quadratic term: β = -0.186, p = 0.871).
+This suggests that the size of the school board, measured by enrollment, has minimal impact on graduation outcomes, with each additional thousand students associated with a non-significant change of 0.00053 percentage points in graduation rates once converted back into real-world units.
 
@@
 @fig-fit shows the relationship between predicted and actual graduation rates.
-The model explains approximately 40.4% of the variance in graduation rates (R² = 0.404), suggesting that while our identified factors are important, there remain other unobserved variables that influence graduation rates.
+The model explains approximately 36.5% of the variance in graduation rates (R² = 0.3648), suggesting that while our identified factors are important, there remain other unobserved variables that influence graduation rates.
 
@@
 ### Data Constraints
+
 1.  Our analysis is cross-sectional, looking at a single academic year (2019-2020)
@@
 ### Methodological Considerations
+
 1.  Our polynomial regression model, while capturing non-linear relationships, may not fully represent more complex interactions between variables
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 # References
-
-https://data.ontario.ca/dataset/school-board-achievements-and-progress/resource/9d81dc27-32ef-4864-94b5-f09950d00c72
-
-https://data.ontario.ca/dataset/school-board-financial-reports-estimates-revised-estimates-and-financial-statements/resource/56598892-c180-4f0d-bde5-63f971d6ef9a
-
-https://data.ontario.ca/dataset/school-information-and-student-demographics/resource/e0e90bd5-d662-401a-a6d2-60d69ac89d14

### 2024-12-04 scripts/05-model.R

@@ -14,27 +14,26 @@ library(glmnet)      # For Ridge and Lasso regression
 
 #### Load data ####
 analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.parquet"))
-colnames(analysis_data) <- gsub(" ", "_", colnames(analysis_data))
 
 #### Data Preprocessing ####
 data <- analysis_data %>%
-  dplyr::select(Four_Year_Graduation_Rate, 
+  dplyr::select(`Four Year Graduation Rate`, 
                 percent_no_degree, 
                 percent_low_income, 
                 percentage_spent_on_facilities,
                 expenses_per_quota,
-                Total_Enrolment)
+                `Total Enrolment`)
 
 # Normalize data
 data_scaled <- data %>%
   mutate(across(everything(), ~ scale(.) %>% as.vector()))
 
 # Prepare data for glmnet
-x <- model.matrix(Four_Year_Graduation_Rate ~ ., data_scaled)[, -1]
-y <- data_scaled$Four_Year_Graduation_Rate
+x <- model.matrix(`Four Year Graduation Rate` ~ ., data_scaled)[, -1]
+y <- data_scaled$`Four Year Graduation Rate`
 
 #### Linear Regression Model ####
-lm_model <- lm(Four_Year_Graduation_Rate ~ ., data = data_scaled)
+lm_model <- lm(`Four Year Graduation Rate` ~ ., data = data_scaled)
 
 #### Ridge Regression ####
 ridge_cv <- cv.glmnet(x, y, alpha = 0)
@@ -47,12 +46,12 @@ lasso_best_lambda <- lasso_cv$lambda.min
 lasso_model <- glmnet(x, y, alpha = 1, lambda = lasso_best_lambda)
 
 #### Polynomial Regression ####
-poly_model <- lm(Four_Year_Graduation_Rate ~ 
+poly_model <- lm(`Four Year Graduation Rate` ~ 
                    poly(percent_no_degree, 2) + 
                    poly(percent_low_income, 2) +
                    poly(percentage_spent_on_facilities, 2) +
                    poly(expenses_per_quota, 2) +
-                   poly(Total_Enrolment, 2),
+                   poly(`Total Enrolment`, 2),
                  data = data_scaled)
 
[…]

### 2024-12-04 paper/paper.qmd
final commit

@@
 summary_stats <- analysis_data %>%
-  summarise(across(c(`Four Year Graduation Rate`, 
+  summarise(across(
+    c(
+      `Four Year Graduation Rate`,
       `Total Expenses`,
@@
       percent_no_degree,
-                     percent_low_income),
+      percent_low_income
+    ),
     list(
-                     Mean = ~mean(., na.rm = TRUE),
-                     SD = ~sd(., na.rm = TRUE),
-                     Min = ~min(., na.rm = TRUE),
-                     Max = ~max(., na.rm = TRUE),
-                     Median = ~median(., na.rm = TRUE)
-                   )))
+      Mean = ~ mean(., na.rm = TRUE),
+      SD = ~ sd(., na.rm = TRUE),
+      Min = ~ min(., na.rm = TRUE),
+      Max = ~ max(., na.rm = TRUE),
+      Median = ~ median(., na.rm = TRUE)
+    )
+  ))
 
@@
 lasso_model <- readRDS(here("models", "lasso_model.rds"))
-
 ```
@@
 Additional file management was facilitated by the here [@here] package.
+Styler [@styler] was used to style the code.
 
@@
   mutate(
-    across(contains("Total Expenses"), ~./1e9),  # Convert to billions
-    across(contains("Total Enrolment"), ~./1000)  # Convert to thousands
+    across(contains("Total Expenses"), ~ . / 1e9), # Convert to billions
+    across(contains("Total Enrolment"), ~ . / 1000) # Convert to thousands
   ) %>%
@@
     names_to = c("Variable", "Statistic"),
-               names_pattern = "(.+)_(.+)") %>%
-  pivot_wider(names_from = Statistic, 
-             values_from = value) %>%
+    names_pattern = "(.+)_(.+)"
+  ) %>%
+  pivot_wider(
+    names_from = Statistic,
+    values_from = value
+  ) %>%
   gt() %>%
@@
   analysis_data |>
-    select(`Board Number`,
+    select(
+      `Board Number`,
       `Board Name`,
@@
       `Four Year Graduation Rate`,
-           `Total Expenses`) |>
+      `Total Expenses`
+    ) |>
     slice(1:5) |>
     mutate(Region = str_replace(Region, "Region ", "")),
-  col.names = c("Board ID",
+  col.names = c(
+    "Board ID",
     "Name",
@@
     "Grad Rate",
-                "Total Expenses ($)"),
+    "Total Expenses ($)"
+  ),
   digits = c(0, 0, 0, 0, 3, 3),
-  booktabs = TRUE)
+  booktabs = TRUE
+)
 ```
+
 ```{r}
@@
 #| tbl-cap: "Sample of First Five Rows from Ontario School Board Dataset (Part 2)"
-
+#| echo: false
+#| warning: false
+#| message: false
 # Second half of columns
@@
   analysis_data |>
-    select(`Expenditure on Facilities`,
+    select(
+      `Expenditure on Facilities`,
       `Total Enrolment`,
@@
       percentage_spent_on_facilities,
-           expenses_per_quota) |>
+      expenses_per_quota
+    ) |>
     slice(1:5),
-  col.names = c("Facility Expenses ($)",
+  col.names = c(
+    "Facility Expenses ($)",
     "Enrolment",
@@
     "Facilities (%)",
-                "$/Student"),
+    "$/Student"
+  ),
   digits = c(3, 0, 3, 3, 3, 3),
-  booktabs = TRUE)
+  booktabs = TRUE
+)
 ```
@@
   geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
-  labs(x = "Four Year Graduation Rate (%)",
-       y = "Number of School Boards") +
+  labs(
+    x = "Four Year Graduation Rate (%)",
+    y = "Number of School Boards"
+  ) +
   theme_minimal()
@@
   geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
-  labs(x = "Percentage Spent on Facilities",
-       y = "Number of School Boards") +
+  labs(
+    x = "Percentage Spent on Facilities",
+    y = "Number of School Boards"
+  ) +
   theme_minimal()
@@
   geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
-  labs(x = "Expenses per Student ($)",
-       y = "Number of School Boards") +
+  labs(
+    x = "Expenses per Student ($)",
+    y = "Number of School Boards"
+  ) +
   theme_minimal()
@@
   geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
-  labs(x = "Percentage of Students Without Parents Holding a Post-Secondary Degree",
-       y = "Number of School Boards") +
+  labs(
+    x = "Percentage of Students Without Parents Holding a Post-Secondary Degree",
+    y = "Number of School Boards"
+  ) +
   theme_minimal()
@@
   geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
-  labs(x = "Percentage of Students in Low Income Families",
-       y = "Number of School Boards") +
+  labs(
+    x = "Percentage of Students in Low Income Families",
+    y = "Number of School Boards"
+  ) +
   theme_minimal()
@@
 #| warning: false
-ggplot(analysis_data, aes(x = `Total Enrolment`/1000)) +
+ggplot(analysis_data, aes(x = `Total Enrolment` / 1000)) +
   geom_histogram(fill = "#2c7bb6", color = "white", alpha = 0.8) +
-  labs(x = "Total Enrolment (Thousands)",
-       y = "Number of School Boards") +
+  labs(
+    x = "Total Enrolment (Thousands)",
+    y = "Number of School Boards"
+  ) +
   theme_minimal()
@@
 data <- analysis_data %>%
-  dplyr::select(`Four Year Graduation Rate`, 
+  dplyr::select(
+    `Four Year Graduation Rate`,
     percent_no_degree,
@@
     expenses_per_quota,
-                `Total Enrolment`)
+    `Total Enrolment`
+  )
 
@@
 #### Polynomial Regression ####
-poly_model <- lm(`Four Year Graduation Rate` ~ 
+poly_model <- lm(
+  `Four Year Graduation Rate` ~
     poly(percent_no_degree, 2) +
@@
     poly(`Total Enrolment`, 2),
-                 data = data_scaled)
+  data = data_scaled
+)
 
@@
   stat_qq_line() +
-  labs(title = "Normal Q-Q Plot",
+  labs(
+    title = "Normal Q-Q Plot",
     x = "Theoretical Quantiles",
-       y = "Sample Quantiles") +
+    y = "Sample Quantiles"
+  ) +
   theme_minimal()
@@
 # Residuals vs Fitted
-ggplot(data.frame(fitted = fitted(poly_model), 
-                 residuals = residuals(poly_model)), 
-       aes(x = fitted, y = residuals)) +
+ggplot(
+  data.frame(
+    fitted = fitted(poly_model),
+    residuals = residuals(poly_model)
+  ),
+  aes(x = fitted, y = residuals)
+) +
   geom_point() +
   geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
-  labs(title = "Residuals vs Fitted Values",
+  labs(
+    title = "Residuals vs Fitted Values",
     x = "Fitted Values",
-       y = "Residuals") +
+    y = "Residuals"
+  ) +
   theme_minimal()
@@
   Metric = c("R-squared", "RMSE"),
-  Value = c(summary(poly_model)$r.squared, 
-            sqrt(mean(residuals(poly_model)^2)))
+  Value = c(
+    summary(poly_model)$r.squared,
+    sqrt(mean(residuals(poly_model)^2))
+  )
 )
@@
 
-The percentage of households without post-secondary degrees shows the strongest relationship with graduation rates (β = -3.627, p = 0.003).
+The percentage of households without post-secondary degrees shows the strongest relationship with graduation rates ($β$ = -3.627, p = 0.003).
 After converting from standardized units back to real-world units, this relationship suggests that for every one percentage point increase in households without post-secondary degrees, graduation rates decrease by approximately 0.136 percentage points, holding other factors constant.
@@
 #| warning: false
-ggplot(data_scaled, aes(x = percentage_spent_on_facilities, 
-                        y = `Four Year Graduation Rate`)) +
+ggplot(data_scaled, aes(
+  x = percentage_spent_on_facilities,
+  y = `Four Year Graduation Rate`
+)) +
   geom_point(alpha = 0.6) +
-  geom_smooth(method = "lm", 
+  geom_smooth(
+    method = "lm",
     formula = y ~ poly(x, 2),
-              se = TRUE) +
-  labs(x = "Percentage Spent on Facilities (Standardized)",
-       y = "Graduation Rate (Standardized)") +
+    se = TRUE
+  ) +
+  labs(
+    x = "Percentage Spent on Facilities (Standardized)",
+    y = "Graduation Rate (Standardized)"
+  ) +
   theme_minimal()
@@
 
-The percentage spent on facilities shows a significant negative linear relationship (β = -2.435, p = 0.023), as illustrated in @fig-facilities.
+The percentage spent on facilities shows a significant negative linear relationship ($β$ = -2.435, p = 0.023), as illustrated in @fig-facilities.
 After converting from standardized units back to real-world units, this suggests that for each additional percentage point of budget allocated to facilities, graduation rates decrease by about 0.197 percentage points, holding other factors constant.
@@
 #| warning: false
-ggplot(data_scaled, aes(x = expenses_per_quota, 
-                        y = `Four Year Graduation Rate`)) +
+ggplot(data_scaled, aes(
+  x = expenses_per_quota,
+  y = `Four Year Graduation Rate`
+)) +
   geom_point(alpha = 0.6) +
-  geom_smooth(method = "lm", 
+  geom_smooth(
+    method = "lm",
     formula = y ~ poly(x, 2),
-              se = TRUE) +
-  labs(x = "Expenses per Student (Standardized)",
-       y = "Graduation Rate (Standardized)") +
+    se = TRUE
+  ) +
+  labs(
+    x = "Expenses per Student (Standardized)",
+    y = "Graduation Rate (Standardized)"
+  ) +
   theme_minimal()
@@
 
-Per-student expenses show a marginally significant quadratic relationship (β = 1.932, p = 0.085), as shown in @fig-expenses.
+Per-student expenses show a marginally significant quadratic relationship ($β$ = 1.932, p = 0.085), as shown in @fig-expenses.
 After converting from standardized units back to real-world units, while not significant at the conventional p < 0.05 level, there appears to be a U-shaped trend in the relationship between per-student spending and graduation rates.
@@
 #| warning: false
-ggplot(data_scaled, aes(x = `Total Enrolment`, 
-                        y = `Four Year Graduation Rate`)) +
+ggplot(data_scaled, aes(
+  x = `Total Enrolment`,
+  y = `Four Year Graduation Rate`
+)) +
   geom_point(alpha = 0.6) +
-  geom_smooth(method = "lm", 
+  geom_smooth(
+    method = "lm",
     formula = y ~ poly(x, 2),
-              se = TRUE) +
-  labs(x = "Total Enrolment (Standardized)",
-       y = "Graduation Rate (Standardized)") +
+    se = TRUE
+  ) +
+  labs(
+    x = "Total Enrolment (Standardized)",
+    y = "Graduation Rate (Standardized)"
+  ) +
   theme_minimal()
@@
 
-As shown in @fig-size, Total Enrolment showed no statistically significant relationship with graduation rates (linear term: β = 0.249, p = 0.842; quadratic term: β = -0.186, p = 0.871).
+As shown in @fig-size, Total Enrolment showed no statistically significant relationship with graduation rates (linear term: $β$ = 0.249, p = 0.842; quadratic term: $β$ = -0.186, p = 0.871).
 This suggests that the size of the school board, measured by enrollment, has minimal impact on graduation outcomes, with each additional thousand students associated with a non-significant change of 0.00053 percentage points in graduation rates once converted back into real-world units.
@@
 #| warning: false
-ggplot(data_scaled, aes(x = fitted(poly_model), 
-                        y = `Four Year Graduation Rate`)) +
+ggplot(data_scaled, aes(
+  x = fitted(poly_model),
+  y = `Four Year Graduation Rate`
+)) +
   geom_point(alpha = 0.6) +
-  geom_abline(intercept = 0, slope = 1, 
-              color = "red", linetype = "dashed") +
-  labs(x = "Predicted Graduation Rate (Standardized)",
-       y = "Actual Graduation Rate (Standardized)") +
+  geom_abline(
+    intercept = 0, slope = 1,
+    color = "red", linetype = "dashed"
+  ) +
+  labs(
+    x = "Predicted Graduation Rate (Standardized)",
+    y = "Actual Graduation Rate (Standardized)"
+  ) +
   theme_minimal()
@@
 
-Our analysis reveals several important insights about the factors influencing graduation rates in Ontario school boards.
-The most striking finding is the strong negative relationship between the percentage of households without post-secondary degrees and graduation rates.
+The key takeaway from the model results is the strong negative relationship between the percentage of households without post-secondary degrees and graduation rates.
 This relationship suggests an intergenerational aspect to educational achievement, where students from communities with lower educational attainment face additional challenges in completing their secondary education.
 
-The complex relationship between per-student expenses and graduation rates is particularly noteworthy.
+The non-linear relationship between per-student expenses and graduation rates is particularly noteworthy.
 The U-shaped relationship we observed suggests that simply increasing funding may not always lead to better outcomes.
@@
 While maintaining adequate facilities is crucial for educational delivery, our findings suggest that boards allocating a larger proportion of their budget to facilities tend to have lower graduation rates.
-This could indicate that:
-
-1.  Some boards might be forced to prioritize urgent infrastructure needs over other educational resources
-2.  Older facilities requiring more maintenance might be concentrated in areas facing other socioeconomic challenges
-3.  There might be an optimal balance between infrastructure investment and direct educational spending
+This could indicate that: some boards might be forced to prioritize urgent infrastructure needs over other educational resources, older facilities requiring more maintenance might be concentrated in areas facing other socioeconomic challenges, or that there might potentially be an optimal balance between infrastructure investment and direct educational spending, and further research in this direction is advised.
 
@@
 
-1.  Our analysis is cross-sectional, looking at a single academic year (2019-2020)
-2.  The COVID-19 pandemic may have influenced both spending patterns and graduation rates during this period
-3.  We lack data on important factors such as:
-    -   Teacher qualifications and experience
-    -   Specific program offerings
-    -   Student-level characteristics
+Our analysis is cross-sectional, looking at a single academic year (2023-2024).
+Therefore, trends existing in this model might not exist in previous years, and vice versa.
+Additionally, due to privacy concerns, lack of granularity in the data and similar reasons, more specific predictors such as the teacher-to-student ratio and student-to-faculty ratio as well as data specifically recording expenditures on libraries are lacking.
+This potentially led to the model not being able to utilise enough, or ideal, predictors in its training.
 
@@
 
-1.  Our polynomial regression model, while capturing non-linear relationships, may not fully represent more complex interactions between variables
-2.  The R-squared value of 0.404 suggests that substantial variation in graduation rates remains unexplained by our model
-3.  The standardization of variables, while necessary for our analysis, makes direct interpretation of effect sizes less intuitive
+Our polynomial regression model, while capturing non-linear relationships, may not fully represent more complex interactions between variables that are of higher degrees.
+The R-squared value of 0.404 suggests that substantial variation in graduation rates remains unexplained by our model, which again is likely due to the lack of data as mentioned above.
 
@@
 
-Several promising avenues for future research emerge from our findings:
-
-1.  **Longitudinal Analysis**: Examining how these relationships evolve over time could provide insights into the long-term effects of different spending patterns and policy changes.
-
-2.
-**Program-Level Analysis**: Investigating specific educational programs and interventions could help identify which approaches are most effective at improving graduation rates, particularly in communities with lower educational attainment.
-
-3.
-**Resource Allocation Optimization**: Developing models to help school boards optimize their resource allocation decisions, particularly regarding the balance between facilities maintenance and other educational spending.
-
-4.
-**Policy Implications**: Studying how provincial funding formulas might be adjusted to better support boards facing particular challenges, such as aging infrastructure or high concentrations of socioeconomic disadvantage.
+Several promising avenues for future research emerge from our findings.
+A longitudinal analysis examining how these relationships evolve over time could provide insights into the long-term effects of different spending patterns and policy changes.
+At a more granular level, investigating specific educational programs and interventions could help identify which approaches are most effective at improving graduation rates, particularly in communities with lower educational attainment.
+Future research could also focus on developing models to help school boards optimize their resource allocation decisions, particularly regarding the balance between facilities maintenance and other educational spending.
+Finally, studying how provincial funding formulas might be adjusted to better support boards facing particular challenges, such as aging infrastructure or high concentrations of socioeconomic disadvantage, would provide valuable policy insights.
 
@@
 The strong relationship between community educational attainment and graduation rates suggests that breaking cycles of educational disadvantage may require interventions that extend beyond the school system itself.
-This might include:
-
-1.  Community-based programs to support adult education
-2.  Enhanced early childhood education initiatives
-3.  Targeted support for communities with historically lower educational attainment
-4.  Integrated approaches that address both educational and broader socioeconomic factors
+This might include community-based programs to support adult education, enhanced early childhood education initiatives, targeted support for communities with historically lower educational attainment, and/or integrated approaches that address both educational and broader socioeconomic factors
 
@@
 
-# Additional data details
-
 # Model details {#sec-model-details}
 
-## Posterior predictive check
+```{r}
+#| label: fig-appendix-cook
+#| fig-cap: "Cook's Distance Plot"
+#| echo: false
+#| message: false
+#| warning: false
+
+plot(cooks.distance(poly_model),
+  type = "h",
+  main = "Cook's Distance",
+  ylab = "Cook's Distance",
+  xlab = "Observation Number"
+)
+abline(
+  h = 4 / length(cooks.distance(poly_model)),
+  col = "red",
+  lty = 2
+)
+```
+
+\FloatBarrier
+
+# Model Diagnostics
+
+@fig-appendix-cook shows Cook's distance for each observation, helping identify potentially influential data points.
+The dashed red line indicates a commonly used threshold of 4/n for identifying influential observations.
+
+This diagnostic helps assess the validity of our model assumptions and identify potential areas of concern in our analysis.
+
+# Data Collection and Sampling Considerations
+
+## Overview of Data Collection Process
+
+The data used in this analysis comes from a complex, multi-source collection process that merges administrative data from Ontario school boards with demographic information and standardized test results.
+Understanding this process is crucial for interpreting our results and identifying potential sources of bias.
+
+## Administrative Data Collection
+
+The Education Financial Information System (EFIS) serves as the primary collection mechanism for financial data.
+School boards submit standardized financial reports through EFIS, which then undergoes several quality control steps.
+These steps include automated validation checks for mathematical accuracy and internal consistency, cross-reference checks against previous years' submissions, manual review by Ministry staff for anomalous patterns, and opportunities for boards to explain unusual variations.
+
+## Sampling Frame and Coverage
+
+While our dataset represents the complete population of public school boards in Ontario, it's important to note that this itself is a subset of all educational institutions in Ontario.
+Specifically, our sampling frame excludes private schools, First Nations schools, and adult education centers.
+This coverage limitation affects the generalizability of our findings.
+
+## Implications for Analysis
+
+These data collection and measurement considerations have several implications for our analysis.
+Our focus on English-language public boards means our findings may not generalize to other educational contexts in Ontario, while financial data, being self-reported, may contain systematic measurement error.
+Furthermore, different variables are measured at different times throughout the school year, potentially introducing temporal misalignment in our cross-sectional analysis, and as stated by Data Ontario [@ontario_cell_reference], data missing from the raw dataset is not missing completely at random (MCAR), which could introduce bias and/or artificially inflated variance in our estimates.
+
+These limitations suggest several directions for future data collection efforts.
+Future work should consider expanding coverage to include private institutions, implementing standardized reporting periods across all variables, developing more robust validation procedures for self-reported financial data, and creating linked datasets that allow for longitudinal analysis.
 

### 2024-12-04 scripts/00-simulate_data.R

@@ -19,21 +19,21 @@ n_rows <- 100
 #### Simulate data ####
 # Generate simulated data for each column
 simulated_data <- tibble(
-  `Board Number` = sprintf("B%03d", 1:n_rows),  # Simulated board numbers (e.g., B001, B002)
-  `Board Name` = paste("Board", 1:n_rows),     # Simulated board names (e.g., Board 1, Board 2)
-  Region = sample(c("Region A", "Region B", "Region C"), n_rows, replace = TRUE),  # Random regions
-  City = sample(c("City X", "City Y", "City Z"), n_rows, replace = TRUE),          # Random cities
-  `Four Year Graduation Rate` = round(runif(n_rows, 0.7, 1), 3),                   # Graduation rates between 0.7 and 1
-  `Total Expenses` = round(runif(n_rows, 1e6, 1e8), 3),                            # Total expenses between $1M and $100M
-  `Expenditure on Facilities` = round(runif(n_rows, 1e5, 1e7), 3),                 # Facility expenses between $100K and $10M
-  `Total Enrolment` = round(runif(n_rows, 100, 5000)),                             # Total enrolment between 100 and 5000 students
-  percent_low_income = round(runif(n_rows, 10, 50), 3),                            # Low-income percentage between 10% and 50%
-  percent_no_degree = round(runif(n_rows, 5, 30), 3)                               # No-degree percentage between 5% and 30%
+  `Board Number` = sprintf("B%03d", 1:n_rows), # Simulated board numbers (e.g., B001, B002)
+  `Board Name` = paste("Board", 1:n_rows), # Simulated board names (e.g., Board 1, Board 2)
+  Region = sample(c("Region A", "Region B", "Region C"), n_rows, replace = TRUE), # Random regions
+  City = sample(c("City X", "City Y", "City Z"), n_rows, replace = TRUE), # Random cities
+  `Four Year Graduation Rate` = round(runif(n_rows, 0.7, 1), 3), # Graduation rates between 0.7 and 1
+  `Total Expenses` = round(runif(n_rows, 1e6, 1e8), 3), # Total expenses between $1M and $100M
+  `Expenditure on Facilities` = round(runif(n_rows, 1e5, 1e7), 3), # Facility expenses between $100K and $10M
+  `Total Enrolment` = round(runif(n_rows, 100, 5000)), # Total enrolment between 100 and 5000 students
[…]

### 2024-12-04 scripts/01-test_simulated_data.R

@@ -17,44 +17,43 @@ simulated_data <- read_parquet(here::here("./data/simulated_data/simulated_data.
 #### Add `expenses_per_quota` column ####
 simulated_data <- simulated_data %>%
   mutate(
-    expenses_per_quota = round(`Total Expenses` / `Total Enrolment`, 3)  # Add and round to 3 decimals
+    expenses_per_quota = round(`Total Expenses` / `Total Enrolment`, 3) # Add and round to 3 decimals
   )
 
 #### Define sanity check function ####
 # Function to perform tests on the dataset
 run_sanity_checks <- function(data, dataset_name) {
   test_that(paste("Sanity Checks for", dataset_name), {
-    
     #### Data structure tests ####
     # Check if the dataset is a dataframe or tibble
     expect_true(is.data.frame(data), info = paste(dataset_name, "is not a dataframe or tibble"))
-    
+
     # Verify that the dataset has an appropriate number of rows
     expect_true(nrow(data) >= 5 && nrow(data) <= 1000,
-                info = paste(dataset_name, "does not have between 5 and 1000 rows")
+      info = paste(dataset_name, "does not have between 5 and 1000 rows")
     )
-    
+
     # Ensure the dataset has exactly 12 columns (new column added)
     expect_equal(ncol(data), 12,
-                 info = paste(dataset_name, "does not have exactly 12 columns")
+      info = paste(dataset_name, "does not have exactly 12 columns")
     )
-    
+
     # Confirm all required columns are present, including the new column
     required_cols <- c(
       "Board Number", "Board Name", "Region", "City",
       "Four Year Graduation Rate", "Total Expenses",
       "Expenditure on Facilities", "Total Enrolment",
-      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities", 
+      "percent_low_income", "percent_no_degree", "percentage_spent_on_facilities",
       "expenses_per_quota"
     )
     expect_true(all(required_cols %in% colnames(data)),
-                info = paste(dataset_name, "is missing required columns")
+      info = paste(dataset_name, "is missing required columns")
     )
-    
+
     #### Data type tests ####
     # Check the types of key variables, including the new column
     expect_type(data$expenses_per_quota, "double")
-    
+
     # Existing type checks
     expect_type(data$`Board Number`, "character")
     expect_type(data$`Board Name`, "character")
@@ -67,61 +66,62 @@ run_sanity_checks <- function(data, dataset_name) {
     expect_type(data$percent_low_income, "double")
[…]

### 2024-12-04 scripts/03-clean_data.R

@@ -17,26 +17,28 @@ raw_data <- read_excel(here::here("./data/raw_data/academic_performance.xlsx"))
 
 # Remove unused variables in the data
 cleaned_data <- raw_data %>%
-  select(-contains("Progress"), 
-         -contains("Credit Accumulation"), 
-         -contains("Results"), 
-         -contains("Language"),
-         -contains("Type"),
-         -contains("Five")) %>%
-  drop_na()  # Cleans out N/A values
+  select(
+    -contains("Progress"),
+    -contains("Credit Accumulation"),
+    -contains("Results"),
+    -contains("Language"),
+    -contains("Type"),
+    -contains("Five")
+  ) %>%
+  drop_na() # Cleans out N/A values
 
 # Read school expenditures data
 raw_data_2 <- read_csv(here::here("./data/raw_data/school_expenditures.csv"))
 
 # Remove unused variables in the data (as well as filler rows/columns in this case)
 cleaned_data_2 <- raw_data_2 %>%
-  slice(c(1, 16, 46)) %>%  # Select specific rows
-  select(9:ncol(raw_data_2))       # Select columns from the 10th onward
+  slice(c(1, 16, 46)) %>% # Select specific rows
+  select(9:ncol(raw_data_2)) # Select columns from the 10th onward
 
 # Transpose data to be in the same format as the first table
 cleaned_data_2 <- t(cleaned_data_2)
 cleaned_data_2 <- as.data.frame(cleaned_data_2)
-colnames(cleaned_data_2) <- c("Board Number", "Total Expenses", "Expenditure on Facilities")  # Assign new column names
+colnames(cleaned_data_2) <- c("Board Number", "Total Expenses", "Expenditure on Facilities") # Assign new column names
 
 # Merge the data from the two tables based on board_number
 merged_data <- cleaned_data %>%
@@ -47,10 +49,12 @@ table3 <- read_excel(here::here("./data/raw_data/education_indicators.xlsx"))
 
 # Select desired variables from the data
 relevant_data <- table3 %>%
-  select(board_number = 1, 
-         enrolment = 22, 
-         percent_low_income = 47, 
-         percent_no_degree = 48)
+  select(
+    board_number = 1,
+    enrolment = 22,
+    percent_low_income = 47,
+    percent_no_degree = 48
+  )
 
 # Set numeric variables to numeric variables
 relevant_data <- relevant_data %>%
[…]

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -18,21 +18,20 @@ analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.par
 # Function to perform tests on the dataset
 run_sanity_checks <- function(data, dataset_name) {
   test_that(paste("Sanity Checks for", dataset_name), {
-    
     #### Data structure tests ####
     # Check if the dataset is a dataframe or tibble
     expect_true(is.data.frame(data), info = paste(dataset_name, "is not a dataframe or tibble"))
-    
+
     # Verify that the dataset has an appropriate number of rows
     expect_true(nrow(data) >= 5 && nrow(data) <= 1000,
-                info = paste(dataset_name, "does not have between 5 and 1000 rows")
+      info = paste(dataset_name, "does not have between 5 and 1000 rows")
     )
-    
+
     # Ensure the dataset has exactly 12 columns
     expect_equal(ncol(data), 12,
-                 info = paste(dataset_name, "does not have exactly 12 columns")
+      info = paste(dataset_name, "does not have exactly 12 columns")
     )
-    
+
     # Confirm all required columns are present
     required_cols <- c(
       "Board Number", "Board Name", "Region", "City",
@@ -42,9 +41,9 @@ run_sanity_checks <- function(data, dataset_name) {
       "expenses_per_quota"
     )
     expect_true(all(required_cols %in% colnames(data)),
-                info = paste(dataset_name, "is missing required columns")
+      info = paste(dataset_name, "is missing required columns")
     )
-    
+
     #### Data type tests ####
     # Check the types of key variables
     expect_type(data$`Board Number`, "character")
@@ -59,67 +58,68 @@ run_sanity_checks <- function(data, dataset_name) {
     expect_type(data$percent_no_degree, "double")
     expect_type(data$percentage_spent_on_facilities, "double")
     expect_type(data$expenses_per_quota, "double")
-    
+
     #### Value range tests ####
     # Verify graduation rates are between 0 and 1
     expect_true(all(data$`Four Year Graduation Rate` >= 0 & data$`Four Year Graduation Rate` <= 1),
-                info = paste(dataset_name, "`Four Year Graduation Rate` values are out of range")
+      info = paste(dataset_name, "`Four Year Graduation Rate` values are out of range")
     )
-    
+
     # Ensure total expenses and facility expenditures are non-negative
     expect_true(all(data$`Total Expenses` >= 0),
-                info = paste(dataset_name, "`Total Expenses` has negative values")
[…]

### 2024-12-04 scripts/05-model.R

@@ -10,19 +10,21 @@ library(tidyverse)
 library(arrow)
 library(broom)
 library(caret)
-library(glmnet)      # For Ridge and Lasso regression
+library(glmnet) # For Ridge and Lasso regression
 
 #### Load data ####
 analysis_data <- read_parquet(here::here("./data/analysis_data/analysis_data.parquet"))
 
 #### Data Preprocessing ####
 data <- analysis_data %>%
-  dplyr::select(`Four Year Graduation Rate`, 
-                percent_no_degree, 
-                percent_low_income, 
-                percentage_spent_on_facilities,
-                expenses_per_quota,
-                `Total Enrolment`)
+  dplyr::select(
+    `Four Year Graduation Rate`,
+    percent_no_degree,
+    percent_low_income,
+    percentage_spent_on_facilities,
+    expenses_per_quota,
+    `Total Enrolment`
+  )
 
 # Normalize data
 data_scaled <- data %>%
@@ -46,13 +48,15 @@ lasso_best_lambda <- lasso_cv$lambda.min
 lasso_model <- glmnet(x, y, alpha = 1, lambda = lasso_best_lambda)
 
 #### Polynomial Regression ####
-poly_model <- lm(`Four Year Graduation Rate` ~ 
-                   poly(percent_no_degree, 2) + 
-                   poly(percent_low_income, 2) +
-                   poly(percentage_spent_on_facilities, 2) +
-                   poly(expenses_per_quota, 2) +
-                   poly(`Total Enrolment`, 2),
-                 data = data_scaled)
+poly_model <- lm(
+  `Four Year Graduation Rate` ~
+    poly(percent_no_degree, 2) +
+    poly(percent_low_income, 2) +
+    poly(percentage_spent_on_facilities, 2) +
+    poly(expenses_per_quota, 2) +
+    poly(`Total Enrolment`, 2),
+  data = data_scaled
+)
 
 #### Model Comparison with MSE ####
 # Predictions

### 2024-12-15 paper/paper.qmd
modificatrions

@@
 ---
-title: "A 1% Increase in Spending on Facilities Decreases Graduation Rates by 0.2%: An Analysis on Ontario School Districts and Related Predictors"
+title: "A 1% Increase in The Proportion of Spending on Facilities Decreases Graduation Rates by 0.2%: An Analysis on Ontario School Districts and Related Predictors"
 author: Andrew Goh
@@
 
-## Feature Selection
+## Feature Engineering, Selection, and Extraction
 
@@
 Similarly, various socioeconomic indicators such as median household income, unemployment rates, and low-income status demonstrated substantial overlap in their variation.
-To address these multicollinearity concerns, we selected representative variables that captured distinct aspects of school board characteristics while minimizing redundancy:
+To address these multicollinearity concerns, we used feature engineering, extraction, and selection to isolate representative variables that capture distinct aspects of school board characteristics while minimizing redundancy:
 
-1.  Per-student expenses was chosen over total expenses to control for board size and provide a standardized measure of financial resources.
-
-2.
+Per-student expenses were calculated from total expenses and total enrolment and was chosen as the analysis metric in the financial dimension over total expenses to control for board size and provide a standardized measure of financial resource usage that can be compared in an isolated environment.
 Percentage spent on facilities was retained as it represents a unique aspect of resource allocation independent of overall spending levels, as well as being reminiscent of the author's initial driving question of whether increased spending on libraries would correlate with better academic performance.
-
-3.
-Among socioeconomic indicators, the percentage of low-income students and parental education levels were selected as they captured different dimensions of socioeconomic status while maintaining relatively low correlation with each other (r \< 0.4).
-
-4.  Total enrolment was included as a control variable to account for potential scale effects that might influence graduation rates independently of other factors.
+Among the (highly intercorrelated) various socioeconomic indicators, the percentage of low-income students and parental education levels were selected as they captured different dimensions of socioeconomic status while maintaining relatively low correlation with each other (r \< 0.4).
+Total enrolment was included as a control variable to account for potential scale effects that might influence graduation rates independently of other factors.
+This, combined with expenses per capita, still include all the information that was present in the original total expenses feature, but in the form of two less-correlated variables that can be individually considered and analysed with an additive model.
 
